---
title: "Collman15v2 Gaba comparisons using Maximum in cube of size ~0.5 micron"
author: "Jesse Leigh Patsolic"
output: 
  html_document:
    keep_md: true
    highlight: pygments
    number_sections: yes
    theme: cerulean
    toc: yes
    toc_depth: 2
---

```{r render, eval=FALSE, echo=FALSE}
require(rmarkdown)
require(meda)
require(rhdf5)
require(rgl)
require(randomForest)
require(plotly)
require(kableExtra)
require(knitr)
require(foreach)
require(doMC)
registerDoMC(4)
rmarkdown::render("Max.Rmd")
system('open Max.html -a "Google Chrome.app"')
```

# MEDA run 

Loading the class labels sent by Forrest.  


The `max` was calculated in a `361x X 361y X 11z` cube about the annotation
centroid. 

```{r cc1}
cl <- read.csv("cleft_class.csv")[-18,]
fmax <- read.csv("collman15v2_Fmax_20171219.csv")
sfmax <- scale(fmax, center = TRUE, scale = TRUE)

loc <- read.csv("locations_collman15v2_Fmax_20171219.csv")
```


```{r lookatit}
table(cl$gaba)
table(cl$postgaba)
```

```{r meda, results = 'hide'}
#ccol <- read.csv("params.csv")[1,-c(1,14)]
ccol <- c('blue', 'blue', 'blue', 'red', 'red', 'red', 'black', 'black', 'green', 'green', 'green', 'green')
set.seed(1234)

L <- list()
L[[1]] <- mlocation(fmax, ccol = ccol)
L[[2]] <- d1heat(sfmax, ccol = ccol)
L[[3]] <- cumvar(sfmax)
L[[4]] <- outliers(sfmax)
L[[5]] <- pairhex(sfmax, ccol = ccol)
L[[6]] <- medacor(fmax, ccol = ccol)
set.seed(1234)
L[[7]] <- hmc(sfmax, ccol = ccol,
              maxDepth = 2, modelNames = "VVV")
#plot(L[[7]])
#plotDend(L[[7]])
#stackM(L[[7]], ccol = ccol, centered = TRUE)
```

Note that we are using `VVE` as the model. 


## mlocation
```{r mlocation, fig.height = 4, fig.width = 8, echo=FALSE}
plot(L[[1]])
```

## d1heat
```{r d1heat, fig.height = 8, fig.width = 8, echo=FALSE}
plot(L[[2]])
```

## cumvar
```{r cumvar, fig.height = 5, fig.width = 5, echo=FALSE}
plot(L[[3]])
```

## outliers
```{r outliers, fig.height = 8, fig.width = 8, echo=FALSE}
plot(L[[4]])
```

## pairhex
```{r pairnex, fig.height = 8, fig.width = 8, echo=FALSE, warning = FALSE}
suppressWarnings(pairhex(sfmax))
```

## correlation
```{r cor, fig.height = 8, fig.width = 8, echo=FALSE}
plot(L[[6]])
```

## dendrogram
```{r dend, fig.height = 4, fig.width = 4, echo=FALSE}
plotDend(L[[7]])
```

## pairs plots
```{r pairs, fig.height = 12, fig.width = 12, echo=FALSE}
c1 <- L[[7]]$dat$labels$col
pairs(sfmax, col = viridis(max(c1))[c1], 
      pch = 19, cex = 0.2)
```

## stacked means
```{r stackM, fig.height = 4, fig.width = 4, echo=FALSE}
stackM(L[[7]], ccol = ccol, centered = TRUE, depth = 2)
```

## cluster means
```{r clusterMeans, fig.height = 4, fig.width = 4, echo=FALSE}
clusterMeans(L[[7]], ccol = ccol)
```

# Set $K = 2$

Here we restrict our hierarchical GMM function to one split only and
then compare with the true gaba/non-gaba class labels.

```{r compare, results = 'hide'}
set.seed(1030)
#set.seed(1234)
h2 <- hmc(sfmax, ccol = ccol, maxDepth = 2, modelNames = 'VVE')
l2 <- h2$dat$labels$col - 1
p0 <- mclust::adjustedRandIndex(l2, cl$gaba)
```

```{r perm}
perms <- foreach(i = 1:1e4, .combine = c) %dopar% {
  set.seed(i*2)
  mclust::adjustedRandIndex(sample(l2), cl$gaba)
}
```

```{r plots, fig.height = 12, fig.width = 12, fig.show = 'hold'}
tmp <- h2$dat$labels$col
pairs(h2$dat$data, 
      col = viridis(max(tmp))[tmp],
      pch = 19, cex = 0.5, main = "colored by prediction")

pairs(h2$dat$data, 
      col = c('darkblue', 'violet')[cl$gaba +1],
      pch = 19, cex = 0.5, main = "colored by truth")
```

```{r pdend, fig.height = 4, fig.width = 4}
plotDend(h2)
stackM(h2, depth = 2, centered = TRUE, ccol = ccol)
```


```{r hist}
hist(perms, xlim = c(min(perms), p0 + 0.25*p0),
     main = "permutation test of ARI values", probability = TRUE)
#hist(perms, probability = TRUE)
abline(v = p0, col = 'red')
```

```{r look2, fig.height = 8, fig.width = 12}
df1 <- data.frame(loc)
df1$gaba <- as.factor(cl$gaba)
df1$classification <- as.factor(tmp -1)
df1$correct_classification <- (df1$classification == df1$gaba)
df1$TN <- (df1$classification == 0 & df1$gaba == 0)
df1$FN <- (df1$classification == 0 & df1$gaba == 1)
df1$FP <- (df1$classification == 1 & df1$gaba == 0)
df1$TP <- (df1$classification == 1 & df1$gaba == 1)
#df1$text <- links

p1 <- ggplot(df1, aes(x = x, y=y,z=z, col = gaba, shape =
                      correct_classification)) + 
  facet_wrap(~ z, ncol = 6) +
  geom_point()

p1
```


### same as above with interactivity
```{r plotly, fig.height = 8, fig.width = 10}
ggplotly(p1)
#p2 <- plot_ly(df1, x = ~x, y = ~y, color = ~gaba,
#			  hoverinfo = 'text',
#				text = ~links)
#p2
#
#htmlwidgets::saveWidget(as_widget(p2), "links.html")
```

### Confusion matrix (classfication on rows)

```{r table}
(ta <- table(classification = tmp-1, df1$gaba))
```

The above table shows that out of `r sum(cl$gaba)` true gaba synapses 
`r ta[1,2]` were mis-classified as non-gaba synapses. 


```{r links, echo = FALSE}
base <- "https://viz.boss.neurodata.io/#!{'layers':{'annotation':{'type':'segmentation'_'source':'boss://https://api.boss.neurodata.io/collman/collman15v2/annotation?'}_'EM25K':{'type':'image'_'source':'boss://https://api.boss.neurodata.io/collman/collman15v2/EM25K?'_'opacity':0.5_'blend':'additive'}_'GABA488':{'type':'image'_'source':'boss://https://api.boss.neurodata.io/collman/collman15v2/GABA488?'_'opacity':0.5_'blend':'additive'_'color':1}_'GAD647':{'type':'image'_'source':'boss://https://api.boss.neurodata.io/collman/collman15v2/GAD647?'_'opacity':0.5_'color':1_'visible':false}_'gephyrin594':{'type':'image'_'source':'boss://https://api.boss.neurodata.io/collman/collman15v2/gephyrin594?'_'opacity':0.5_'color':2_'visible':false}_'GS594':{'type':'image'_'source':'boss://https://api.boss.neurodata.io/collman/collman15v2/GS594?'_'opacity':0.5_'color':3_'visible':false}_'MBP488':{'type':'image'_'source':'boss://https://api.boss.neurodata.io/collman/collman15v2/MBP488?'_'opacity':0.5_'color':4_'visible':false}_'NR1594':{'type':'image'_'source':'boss://https://api.boss.neurodata.io/collman/collman15v2/NR1594?'_'opacity':0.5_'color':5_'visible':false}_'PSD95_488':{'type':'image'_'source':'boss://https://api.boss.neurodata.io/collman/collman15v2/PSD95_488?'_'opacity':0.5_'color':6_'visible':false}_'Synapsin647':{'type':'image'_'source':'boss://https://api.boss.neurodata.io/collman/collman15v2/Synapsin647?'_'opacity':0.5_'visible':false}_'VGluT1_647':{'type':'image'_'source':'boss://https://api.boss.neurodata.io/collman/collman15v2/VGluT1_647?'_'opacity':0.5_'color':1_'visible':false}}_'navigation':{'pose':{'position':{'voxelSize':[2.240000009536743_2.240000009536743_70]_'voxelCoordinates':[%s_%s_%s]}}_'zoomFactor':2.25}}"

links <- foreach(i = 1:dim(loc)[1]) %do% {
	sprintf(base,loc$x[i], loc$y[i], loc$z[i])
}

mdlinks <- foreach(i = 1:length(links)) %do% {
	sprintf("[%d, %d, %d](%s)", loc$x[i], loc$y[i], loc$z[i],links[[i]])
}

#fooFN <- file("FN_Gaba.txt")
#writeLines(Reduce(c,
#                  links[df1$FN]
#                  ), fooFN)
#close(fooFN)
#
#fooFP <- file("FP_Gaba.txt")
#writeLines(Reduce(c,
#                  links[df1$FP]
#                  ), fooFP)
#close(fooFP)
#
#write.csv(loc[df1$FN,c(3,2,1)], file = "GabaFN_locXYZ.csv", row.names = FALSE)
#write.csv(loc[df1$FP,c(3,2,1)], file = "GabaFP_locXYZ.csv", row.names = FALSE)
#write.csv(loc[df1$TN,c(3,2,1)], file = "GabaTN_locXYZ.csv", row.names = FALSE)
#write.csv(loc[df1$TP,c(3,2,1)], file = "GabaTP_locXYZ.csv", row.names = FALSE)
```


# NDVIZ Links

Note that the centroids are caluclated from the annotations which means
that the centorid might not be in an area with annotated pixels. 

```{r linkTables, echo = FALSE}
lFN <- data.frame(FalseNegatives = as.character(mdlinks[df1$FN]))
lFP <- data.frame(FalsePositives = as.character(mdlinks[df1$FP]))
lTN <- data.frame(TrueNegatives  = as.character(mdlinks[df1$TN]))
lTP <- data.frame(TruePositives  = as.character(mdlinks[df1$TP]))
```

## False Negatives 
```{r tFN, echo = FALSE}
kable(lFN, format = 'markdown', escape = FALSE)
```

## False Positives 
```{r tFP, echo = FALSE}
kable(lFP, format = 'markdown', escape = FALSE)
```

## True Negatives
```{r tTN, echo = FALSE}
kable(lTN, format = 'markdown', escape = FALSE)
```

## True Positives
```{r tTP, echo = FALSE}
kable(lTP, format = 'markdown', escape = FALSE)
```


# RF

```{r rf}
set.seed(317)
rfdat <- data.table(gaba = as.factor(cl$gaba), sfmax)
rf1 <- randomForest(gaba ~ ., data = rfdat)

#set.seed(317)
set.seed(2^13)
train <- sample(nrow(rfdat), 100)
test <- setdiff(1:nrow(rfdat), train)
table(rfdat[test,]$gaba)

rf2 <- randomForest(gaba ~ ., data = rfdat[train,], importance = TRUE)
print(rf2)

rf2.pred <- predict(rf2, newdata = rfdat[test,-c('gaba')])

table(rf2.pred, rfdat[test,]$gaba)
varImpPlot(rf2)
```


