---
title: "Synapse Clustering: Feature Exploration"
author: "JLP"
date: '`r Sys.Date()`'
output:
  html_document:
    fig_caption: yes
    fig_height: 5
    fig_width: 5
    highlight: pygments
    keep_md: yes
    number_sections: yes
    theme: cerulean
    toc: yes
    toc_depth: 3
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
---
```{r render, eval=FALSE, echo=FALSE}
require(rmarkdown)
rm(list=ls()); 
rmarkdown::render(grep("FeatureExploration.Rmd", dir(), value=TRUE))
system('open FeatureExploration.html')
```
```{r setup,include=FALSE,results='asis',message=FALSE,warning=FALSE}
### Library calls here.
require(Matrix)
require(colorRamps)
require(corrplot)
require(gplots)
require(dplyr)
require(reshape2)
require(ggplot2)
require(gridExtra)
require(lattice)
require(parallel)
require(data.table)
require(fpc)
require(repr)
require(mclust)
require(doMC)
require(rgl)
require(rglwidget)

source("./bic.r")
source("./clusterFraction.r")
source("./getElbows.r")
#source("http://www.cis.jhu.edu/~parky/Synapse/getElbows.R")
registerDoMC(6)

library(rmarkdown)
library(knitr)
```
```{r knitr-setup, include=FALSE, results='asis'}
### The following options and figure numbering functions
### were setup by Youngser Park
knitr::opts_chunk$set(cache=TRUE, autodep=TRUE)
dep_auto() # figure out dependencies automatically
opts_chunk$set(cache=FALSE,echo=TRUE,warning=FALSE,message=FALSE,comment="#",fig.path='../Figures/FeatureExploration_figure/',dpi=227,dev=c('png','pdf'))

opts_knit$set(aliases=c(h='fig.height', w='fig.width', cap='fig.cap', scap='fig.scap'))                                                                               
opts_knit$set(eval.after = c('fig.cap','fig.scap'))                                                                            
knit_hooks$set(document = function(x) {
  gsub('(\\\\end\\{knitrout\\}[\n]+)', '\\1\\\\noindent ', x)                                                   })
#opts_knit$set(animation.fun = hook_scianimator)

knit_hooks$set(plot = function(x, options) {
       paste('<figure><img src="',
             opts_knit$get('base.url'), paste(x, collapse = '.'),
             '"><figcaption>', options$fig.cap, '</figcaption></figure>',
             sep = '')
 })
 
 fn = local({
   i = 0
   function(x) {
     i <<- i + 1
     paste('Figure ', i, ': ', x, sep = '')
   }
 })

 fig <- local({
    i <- 0
    ref <- list()
    list(
        cap=function(refName, text) {
            i <<- i + 1
            ref[[refName]] <<- i
            paste("<b>Figure ", i, ": ", text, "</b><br><br>", sep="")
        },
        ref=function(refName) {
            ref[[refName]]
        })
})
```

[Homepage](http://docs.neurodata.io/synaptome-stats/)  
The formatted source code for this file is [here](https://github.com/neurodata/synaptome-stats/blob/gh-pages/Code/FeatureExploration.Rmd).  
And a [raw version here](https://raw.githubusercontent.com/neurodata/synaptome-stats/gh-pages/Code/FeatureExploration.Rmd).    
Previous work by Youngser Park can be found [here](http://www.cis.jhu.edu/~parky/Synapse/synapse.html).  


# Introduction

Following from previous pages, this page will focus on filtering the data before clustering 
to explore if filtering improves the outcome of clustering.

# Data 
Here we read in the data and select a random half of it for exploration. 

```{r cc-data, eval=TRUE}
featFull <- fread("../data/synapsinR_7thA.tif.Pivots.txt.2011Features.txt",showProgress=FALSE)

### Setting a seed and creating an index vector
### to select half of the data
set.seed(2^10)
half1 <- sample(dim(featFull)[1],dim(featFull)[1]/2)
half2 <- setdiff(1:dim(featFull)[1],half1)

feat <- featFull[half1,]
dim(feat)

## Setting the channel names
channel <- c('Synap_1','Synap_2','VGlut1_t1','VGlut1_t2','VGlut2','Vglut3',
              'psd','glur2','nmdar1','nr2b','gad','VGAT',
              'PV','Gephyr','GABAR1','GABABR','CR1','5HT1A',
              'NOS','TH','VACht','Synapo','tubuli','DAPI')

## Setting the channel types
channel.type <- c('ex.pre','ex.pre','ex.pre','ex.pre','ex.pre','in.pre.small',
                  'ex.post','ex.post','ex.post','ex.post','in.pre','in.pre',
                  'in.pre','in.post','in.post','in.post','in.pre.small','other',
                  'ex.post','other','other','ex.post','none','none')

nchannel <- length(channel)
nfeat <- ncol(feat) / nchannel

## Createing factor variables for channel and channel type sorted properly
ffchannel <- (factor(channel.type,
    levels= c("ex.pre","ex.post","in.pre","in.post","in.pre.small","other","none")
    ))
fchannel <- as.numeric(factor(channel.type,
    levels= c("ex.pre","ex.post","in.pre","in.post","in.pre.small","other","none")
    ))
ford <- order(fchannel)


## Setting up colors for channel types
Syncol <- c("#197300","#5ed155","#660000","#cc0000","#ff9933","mediumblue","gold")
ccol <- Syncol[fchannel]

exType <- factor(c(rep("ex",11),rep("in",6),rep("other",7)),ordered=TRUE)
exCol<-exType;levels(exCol) <- c("#197300","#990000","mediumblue");
exCol <- as.character(exCol)

fname <- as.vector(sapply(channel,function(x) paste0(x,paste0("F",0:5))))
names(feat) <- fname
fcol <- rep(ccol, each=6)
mycol <- colorpanel(100, "purple", "black", "green")
mycol2 <- matlab.like(nchannel)
```

## Data transformations

```{r cc-datatrans, eval=TRUE}
f <- lapply(1:6,function(x){seq(x,ncol(feat),by=nfeat)})
featF <- lapply(f,function(x){subset(feat,select=x)})

featF0 <- featF[[1]]
f01e3 <- 1e3*data.table(apply(X=featF0, 2, function(x){((x-min(x))/(max(x)-min(x)))}))

fs <- f01e3

### Taking log_10 on data with 0's removed
ans <- apply(featF0, 1, function(row){ any(row == 0)})

logF0 <- log10(featF0[!ans,])
slogF0 <- logF0[,lapply(.SD,scale, center=TRUE,scale=TRUE)]
```

We now have the following data sets:

- `featF0`: The feature vector looking only at the integrated brightness features.
- `fs`:  The feature vector scaled between $[0,1000]$.
- `logF0`: The feature vector, with 0's removed, then $log_{10}$ is applied. 
- `slogF0`: The feature vector, with 0's removed, then $log_{10}$, then scaled by
  subtracting the mean and dividing by the sample standard deviation.


# Feature Exploration

## Synapsin1 Vs. Synapsin2 for all features

```{r cc-syn,w=9,h=6,fig.cap=fig$cap("Syn1v2F","Scatter plots of Synap1F0 and Synap2F0 on linear and log scale for all features.")}
synF <- feat[, grep("Synap_", names(feat)),with=FALSE]
ans1 <- apply(synF, 1, function(row){ any(row == 0)})
lsynF <- synF[!ans1,lapply(.SD,function(x){scale(log10(x),center=TRUE,scale=TRUE)})]

print(paste("removed", sum(ans1), "zero entries"))

gg1 <- ggplot(data=synF[,.(Synap_1F0,Synap_2F0)],aes(x=Synap_1F0,y=Synap_2F0)) +   
        geom_point(pch='.',alpha=0.2) + 
        geom_hex(bins=100) +
        geom_smooth(colour='red', alpha=0.7)+
        ggtitle("Untransformed Features: F0")
        
gg2 <- ggplot(data=lsynF[,.(Synap_1F0,Synap_2F0)],aes(x=Synap_1F0,y=Synap_2F0)) +
        geom_point(pch='.',alpha=0.2) +
        geom_hex(bins=100) +
        geom_smooth(colour='red', alpha=0.7)+
        ggtitle("Scaled Logged Features: F0")

gg3 <- ggplot(data=synF[,.(Synap_1F1,Synap_2F1)],aes(x=Synap_1F1,y=Synap_2F1)) +   
        geom_point(pch='.',alpha=0.2) + 
        geom_hex(bins=100) +
        geom_smooth(colour='red', alpha=0.7)+
        ggtitle("Untransformed Features: F1")
        
gg4 <- ggplot(data=lsynF[,.(Synap_1F1,Synap_2F1)],aes(x=Synap_1F1,y=Synap_2F1)) +   
        geom_point(pch='.',alpha=0.2) +
        geom_hex(bins=100) +
        geom_smooth(colour='red', alpha=0.7)+
        ggtitle("Scaled Logged Features: F1")

gg5 <- ggplot(data=synF[,.(Synap_1F2,Synap_2F2)],aes(x=Synap_1F2,y=Synap_2F2)) +   
        geom_point(pch='.',alpha=0.2) + 
        geom_hex(bins=100) +
        geom_smooth(colour='red', alpha=0.7)+
        ggtitle("Untransformed Features: F2")
        
gg6 <- ggplot(data=lsynF[,.(Synap_1F2,Synap_2F2)],aes(x=Synap_1F2,y=Synap_2F2)) +   
        geom_point(pch='.',alpha=0.2) +
        geom_hex(bins=100) +
        geom_smooth(colour='red', alpha=0.7)+
        ggtitle("Scaled Logged Features: F2")

gg7 <- ggplot(data=synF[,.(Synap_1F3,Synap_2F3)],aes(x=Synap_1F3,y=Synap_2F3)) +   
        geom_point(pch='.',alpha=0.2) + 
        geom_hex(bins=100) +
        geom_smooth(colour='red', alpha=0.7)+
        ggtitle("Untransformed Features: F3")
        
gg8 <- ggplot(data=lsynF[,.(Synap_1F3,Synap_2F3)],aes(x=Synap_1F3,y=Synap_2F3)) +   
        geom_point(pch='.',alpha=0.2) +
        geom_hex(bins=100) +
        geom_smooth(colour='red', alpha=0.7)+
        ggtitle("Scaled Logged Features: F3")

gg09 <- ggplot(data=synF[,.(Synap_1F4,Synap_2F4)],aes(x=Synap_1F4,y=Synap_2F4)) +   
        geom_point(pch='.',alpha=0.2) + 
        geom_hex(bins=100) +
        geom_smooth(colour='red', alpha=0.7)+
        ggtitle("Untransformed Features: F4")
        
gg10 <- ggplot(data=lsynF[,.(Synap_1F4,Synap_2F4)],aes(x=Synap_1F4,y=Synap_2F4)) +   
        geom_point(pch='.',alpha=0.2) +
        geom_hex(bins=100) +
        geom_smooth(colour='red', alpha=0.7)+
        ggtitle("Scaled Logged Features: F4")

gg11 <- ggplot(data=synF[,.(Synap_1F5,Synap_2F5)],aes(x=Synap_1F5,y=Synap_2F5)) +   
        geom_point(pch='.',alpha=0.2,position='jitter') + 
        geom_hex(bins=100) +
        geom_smooth(colour='red', alpha=0.7)+
        ggtitle("Untransformed Features: F5")
        
gg12 <- ggplot(data=lsynF[,.(Synap_1F5,Synap_2F5)],aes(x=Synap_1F5,y=Synap_2F5)) +   
        geom_point(pch='.',alpha=0.2, position='jitter') +
        geom_hex(bins=100) +
        geom_smooth(colour='red', alpha=0.7)+
        ggtitle("Scaled Logged Features: F5")
```


```{r cc-F1-5,w=8.5,h=23,fig.cap=fig$cap("synF0-5","Scatter plots of Synapsin1 and Synapsin2 on linear and log scale.")}
grid.arrange(gg1,gg2,gg3,gg4,gg5,gg6,gg7,gg8,gg09,gg10, gg11, gg12, ncol=2)
```

## VGlut1_t1 Vs. VGlut1_t2 for all features

```{r cc-vglut,w=9,h=6,fig.cap=fig$cap("Vglut1v2F","Scatter plots of VGlut1_t1F0 and VGlut1_t2F0 on linear and log scale for all features.")}
vglutF <- feat[,grep("VGlut", names(feat)),with=FALSE]
ans2 <- apply(vglutF,1,function(row){ any(row == 0)})
lvglutF <- vglutF[!ans2,lapply(.SD, function(x){scale(log10(x+1),center=TRUE,scale=TRUE)})]

print(paste("removed", sum(ans2), "zero entries"))

gg13 <- ggplot(data=vglutF[,.(VGlut1_t1F0,VGlut1_t2F0)],aes(x=VGlut1_t1F0,y=VGlut1_t2F0)) +   
        geom_point(pch='.',alpha=0.2) + 
        geom_hex(bins=100) +
        geom_smooth(colour='red', alpha=0.7)+
        ggtitle("Untransformed Features: F0")
        
gg14 <- ggplot(data=lvglutF[,.(VGlut1_t1F0,VGlut1_t2F0)],aes(x=VGlut1_t1F0,y=VGlut1_t2F0)) +   
        geom_point(pch='.',alpha=0.2) +
        geom_hex(bins=100) +
        geom_smooth(colour='red', alpha=0.7)+
        ggtitle("Scaled Logged Features: F0")

gg15 <- ggplot(data=vglutF[,.(VGlut1_t1F1,VGlut1_t2F1)],aes(x=VGlut1_t1F1,y=VGlut1_t2F1)) +   
        geom_point(pch='.',alpha=0.2) + 
        geom_hex(bins=100) +
        geom_smooth(colour='red', alpha=0.7)+
        ggtitle("Untransformed Features: F1")
        
gg16 <- ggplot(data=lvglutF[,.(VGlut1_t1F1,VGlut1_t2F1)],aes(x=VGlut1_t1F1,y=VGlut1_t2F1)) +   
        geom_point(pch='.',alpha=0.2) +
        geom_hex(bins=100) +
        geom_smooth(colour='red', alpha=0.7)+
        ggtitle("Scaled Logged Features: F1")

gg17 <- ggplot(data=vglutF[,.(VGlut1_t1F2,VGlut1_t2F2)],aes(x=VGlut1_t1F2,y=VGlut1_t2F2)) +   
        geom_point(pch='.',alpha=0.2) + 
        geom_hex(bins=100) +
        geom_smooth(colour='red', alpha=0.7)+
        ggtitle("Untransformed Features: F2")
        
gg18 <- ggplot(data=lvglutF[,.(VGlut1_t1F2,VGlut1_t2F2)],aes(x=VGlut1_t1F2,y=VGlut1_t2F2)) +   
        geom_point(pch='.',alpha=0.2) +
        geom_hex(bins=100) +
        geom_smooth(colour='red', alpha=0.7)+
        ggtitle("Scaled Logged Features: F2")

gg19 <- ggplot(data=vglutF[,.(VGlut1_t1F3,VGlut1_t2F3)],aes(x=VGlut1_t1F3,y=VGlut1_t2F3)) +   
        geom_point(pch='.',alpha=0.2) + 
        geom_hex(bins=100) +
        geom_smooth(colour='red', alpha=0.7)+
        ggtitle("Untransformed Features: F3")
        
gg20 <- ggplot(data=lvglutF[,.(VGlut1_t1F3,VGlut1_t2F3)],aes(x=VGlut1_t1F3,y=VGlut1_t2F3)) +   
        geom_point(pch='.',alpha=0.2) +
        geom_hex(bins=100) +
        geom_smooth(colour='red', alpha=0.7)+
        ggtitle("Scaled Logged Features: F3")

gg21 <- ggplot(data=vglutF[,.(VGlut1_t1F4,VGlut1_t2F4)],aes(x=VGlut1_t1F4,y=VGlut1_t2F4)) +   
        geom_point(pch='.',alpha=0.2) + 
        geom_hex(bins=100) +
        geom_smooth(colour='red', alpha=0.7)+
        ggtitle("Untransformed Features: F4")
        
gg22 <- ggplot(data=lvglutF[,.(VGlut1_t1F4,VGlut1_t2F4)],aes(x=VGlut1_t1F4,y=VGlut1_t2F4)) +   
        geom_point(pch='.',alpha=0.2) +
        geom_hex(bins=100) +
        geom_smooth(colour='red', alpha=0.7)+
        ggtitle("Scaled Logged Features: F4")

gg23 <- ggplot(data=vglutF[,.(VGlut1_t1F5,VGlut1_t2F5)],aes(x=VGlut1_t1F5,y=VGlut1_t2F5)) +   
        geom_point(pch='.',alpha=0.2, position='jitter') + 
        geom_hex(bins=100) +
        geom_smooth(colour='red', alpha=0.7)+
        ggtitle("Untransformed Features: F5")
        
gg24 <- ggplot(data=lvglutF[,.(VGlut1_t1F5,VGlut1_t2F5)],aes(x=VGlut1_t1F5,y=VGlut1_t2F5)) +   
        geom_point(pch='.',alpha=0.2, position='jitter') +
        geom_hex(bins=100) +
        geom_smooth(colour='red', alpha=0.7)+
        ggtitle("Scaled Logged Features: F5")
```

```{r cc-vglutF1-5,w=8.5,h=23,fig.cap=fig$cap("vglutF0-5","Scatter plots of VGlut1_t1 and VGlut1_t2 on linear and log scale for all features.")}
grid.arrange(gg13, gg14, gg15, gg16, gg17, gg18, gg19, gg20,gg21,gg22,gg23,gg24, ncol=2)
```

## KDE plots of chosen transformation/feature pair

```{r cc-kde-data}
synF <- feat[, grep("Synap_", names(feat)),with=FALSE]
lsynF <- synF[,lapply(.SD,function(x){scale(log10(x+1),center=TRUE,scale=TRUE)})]
names(synF) <- paste0(names(synF), "_linear")
names(lsynF) <- paste0(names(lsynF), "_logscale")

vglutF <- feat[,grep("VGlut1_t",names(feat)),with=FALSE]
lvglutF <- vglutF[,lapply(.SD,function(x){scale(log10(x+1),center=TRUE,scale=TRUE)})]
names(vglutF) <- paste0(names(vglutF), "_linear")
names(lvglutF) <- paste0(names(lvglutF),"_logscale")
```

```{r cc-kde-syn, w=14,h=8,fig.cap=fig$cap("synKDE", "KDE for Synapsin1 and Synapsin2 accross all features.")}
df1 <- melt(as.matrix(cbind(synF,lsynF)))
levels(df1$Var2)<-levels(df1$Var2)[c(1:6,13:18,7:12,19:24)]
ggplot(data=df1,aes(x=value,y=..density..,group=as.factor(Var2),colour=Var2)) + 
    geom_density() + 
    facet_wrap( ~ Var2,scales='free',ncol=6)
```

```{r cc-kde-vglut, w=14,h=8,fig.cap=fig$cap("vglutKDE", "KDE for VGlut1_t1 and VGlut1_t2 accross all features.")}
df2 <- melt(as.matrix(cbind(vglutF,lvglutF)))
levels(df2$Var2)<-levels(df2$Var2)[c(1:6,13:18,7:12,19:24)]
ggplot(data=df2,aes(x=value,y=..density..,group=as.factor(Var2),colour=Var2)) + 
    geom_density() + 
    facet_wrap( ~ Var2,scales='free', ncol=6)
```




<footer>
<p> [Back to Top][Introduction]</p>
</footer>

