---
title: "Synaptome Stats: Marker Exploration"
author: "Jesse Leigh Patsolic"
date: '`r Sys.Date()`'
output:
  html_document:
    fig_caption: yes
    fig_height: 5
    fig_width: 5
    fig_retina: 2.7
    highlight: pygments
    keep_md: yes
    number_sections: yes
    theme: cerulean
    toc: yes
    toc_depth: 3
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
---
```{r render, eval=FALSE, echo=FALSE}
require(rmarkdown)
rm(list=ls()); 
system.time(rmarkdown::render(grep("MarkerExploration.Rmd", dir(), value=TRUE)))
system('open MarkerExploration.html')
```
```{r setup,include=FALSE,results='asis',message=FALSE}
### Library calls here.
library(rmarkdown)
library(knitr)
require(ggplot2)
require(gridExtra)
require(gplots)
require(lattice)
require(hexbin)
require(data.table)
library(deldir) # for voronoi
require(dplyr)
require(energy)
require(lattice)
require(colorRamps)
require(corrplot)
require(rgl)
require(rglwidget)
require(doMC); registerDoMC(6)
require(foreach)
require(MASS)

source("kmpp.r")
source("clusterFraction.r")
source("http://www.cis.jhu.edu/~parky/Synapse/getElbows.R")
```

```{r knitrOpts, include=FALSE, results='asis', message=FALSE, warning=FALSE}
### The following options and figure numbering functions
### were setup by Youngser Park
knitr::opts_chunk$set(cache=TRUE, autodep=TRUE)
dep_auto() # figure out dependencies automatically
opts_chunk$set(cache=FALSE, echo=TRUE, 
               opts_chunk$set(fig.path = '../Figures/MarkerExploration_figure/'),
               warning=FALSE, message=FALSE, 
               comment="#",
               dpi=227,dev=c('png','pdf'))

opts_knit$set(aliases = c(h = 'fig.height', w = 'fig.width', 
                          cap='fig.cap', scap='figscap'))

opts_knit$set(eval.after = c('fig.cap','fig.scap'))

knit_hooks$set(document = function(x) {
  gsub('(\\\\end\\{knitrout\\}[\n]+)', '\\1\\\\noindent ', x)
  }
)
#opts_knit$set(animation.fun = hook_scianimator)

knit_hooks$set(plot = function(x, options) {
  paste('<figure><img src="',
        opts_knit$get('base.url'), 
        paste(x, collapse = '.'),
             '"><figcaption>', options$fig.cap, '</figcaption></figure>',
             sep = '')
 })
 
 fn = local({
   i = 0
   function(x) {
     i <<- i + 1
     paste('Figure ', i, ': ', x, sep = '')
   }
 })

 fig <- local({
    i <- 0
    ref <- list()
    list(
        cap=function(refName, text) {
            i <<- i + 1
            ref[[refName]] <<- i
            paste("<b>Figure ", i, ": ", text, "</b><br><br>", sep="")
        },
        ref=function(refName) {
            ref[[refName]]
        })
})
```

[Homepage](http://docs.neurodata.io/synaptome-stats/)  
The formatted source code for this file is [here](https://github.com/neurodata/synaptome-stats/blob/gh-pages/Code/MarkerExploration.Rmd).  
And a raw version [here](https://raw.githubusercontent.com/neurodata/synaptome-stats/gh-pages/Code/MarkerExploration.Rmd).    
Previous work by Youngser Park can be found [here](http://www.cis.jhu.edu/~parky/Synapse/synapse.html).  

# Data 

Here we read in the data and select a random half of it for exploration. 

```{r cc-data, eval=TRUE}
featFull <- fread("../data/synapsinR_7thA.tif.Pivots.txt.2011Features.txt",showProgress=FALSE)

### Setting a seed and creating an index vector
### to select half of the data
set.seed(2^10)
half1 <- sample(dim(featFull)[1],dim(featFull)[1]/2)
half2 <- setdiff(1:dim(featFull)[1],half1)

feat <- featFull[half1,]
dim(feat)

## Setting the channel names
channel <- c('Synap_1','Synap_2','VGlut1_t1','VGlut1_t2','VGlut2','Vglut3',
              'psd','glur2','nmdar1','nr2b','gad','VGAT',
              'PV','Gephyr','GABAR1','GABABR','CR1','5HT1A',
              'NOS','TH','VACht','Synapo','tubuli','DAPI')

## Setting the channel types
channel.type <- c('ex.pre','ex.pre','ex.pre','ex.pre','ex.pre','in.pre.small',
                  'ex.post','ex.post','ex.post','ex.post','in.pre','in.pre',
                  'in.pre','in.post','in.post','in.post','in.pre.small','other',
                  'ex.post','other','other','ex.post','none','none')

channel.type2 <- c('ex.pre','ex.pre','ex.pre','ex.pre','ex.pre','other',
                   'ex.post','ex.post','ex.post','ex.post','in.pre','in.pre',
                   'in.pre','in.post','in.post','in.post','other','other',
                   'ex.post','other','other','ex.post','other','other')
nchannel <- length(channel)
nfeat <- ncol(feat) / nchannel

## Createing factor variables for channel and channel type sorted properly
ffchannel <- (factor(channel.type,
    levels= c("ex.pre","ex.post","in.pre","in.post","in.pre.small","other","none")
    ))
fchannel <- as.numeric(factor(channel.type,
    levels= c("ex.pre","ex.post","in.pre","in.post","in.pre.small","other","none")
    ))
ford <- order(fchannel)


## Setting up colors for channel types
Syncol <-  c("#197300","#5ed155","#660000","#cc0000","#ff9933","#0000cd","#ffd700")
Syncol3 <- c("#197300","#197300","#cc0000","#cc0000","#0000cd","#0000cd","#0000cd")
ccol <- Syncol[fchannel]
ccol3 <- Syncol3[fchannel]

exType <- factor(c(rep("ex",11),rep("in",6),rep("other",7)),ordered=TRUE)
exCol<-exType;levels(exCol) <- c("#197300","#990000","#0000cd");
exCol <- as.character(exCol)

fname <- as.vector(sapply(channel,function(x) paste0(x,paste0("F",0:5))))
names(feat) <- fname
fcol <- rep(ccol, each=6)
mycol <- colorpanel(100, "purple", "black", "green")
mycol2 <- matlab.like(nchannel)
mycol3 <- colorpanel(100, "blue","white","red")
```

## Data transformations

```{r cc-datatrans, eval=TRUE}
f <- lapply(1:6,function(x){seq(x,ncol(feat),by=nfeat)})
featF <- lapply(f,function(x){subset(feat,select=x)})

featF0 <- featF[[1]]
f01e3 <- 1e3*data.table(apply(X=featF0,2,function(x){((x-min(x))/(max(x)-min(x)))}))

fs <- f01e3

### Taking log_10 on data + 1.
log1f <- log10(featF0 + 1)
slog1f <- data.table(scale(log1f, center=TRUE,scale=TRUE))
```

We now have the following data sets:

- `featF0`: The feature vector looking only at the integrated brightness features.
- `fs`:  The feature vector scaled between $[0,1000]$.
- `logf1`: The feature vector, plus one, then $log_{10}$ is applied. 
- `slog1f`: The feature vector, plus one, $log_{10}$, then scaled by
  subtracting the mean and dividing by the sample standard deviation.

# Marker Exploration 

## Correlation Matrix of markers 

```{r cc_corRawF0,w=7,h=7, eval=TRUE,fig.cap=fig$cap("corr","Correlation on untransformed F0 data, reordered by synapse type.")}
corrf <- cor(featF[[1]])[ford,ford]
corrplot(corrf,method="color",tl.col=ccol[ford], tl.cex=0.8)
corLog <- cor(slog1f)
```

## Energy Matrix: Distance Correlation T-test

```{r cc-dcov,eval=TRUE,echo=TRUE}
set.seed(317)
sam1 <- sample(dim(featF0)[1], 1e3)
tmp <- as.data.frame((featF0[sam1,]))

set.seed(331)
combcols <- t(combn(24,2))

dc <- foreach(i = 1:dim(combcols)[1]) %dopar% {
       #dcov.test(x=tmp[,combcols[i,1]],y=tmp[,combcols[i,2]])
       dcor.ttest(x=tmp[,combcols[i,1]],y=tmp[,combcols[i,2]])
       }

ms <- matrix(as.numeric(0),24,24)
mp <- matrix(as.numeric(0),24,24)

for(i in 1:length(dc)){
    ms[combcols[i,1],combcols[i,2]] <- dc[[i]]$statistic
    ms[combcols[i,2],combcols[i,1]] <- dc[[i]]$statistic
    mp[combcols[i,1],combcols[i,2]] <- dc[[i]]$p.val
    mp[combcols[i,2],combcols[i,1]] <- dc[[i]]$p.val
}

rownames(ms) <- colnames(featF0)
rownames(mp) <- colnames(featF0)
colnames(ms) <- colnames(featF0)
colnames(mp) <- colnames(featF0)

diag(ms) <- as.numeric(0)
diag(mp) <- as.numeric(1)
```

```{r cc-energyplot,h=7,w=14,fig.cap=fig$cap("enplt","F0: Distance correlation t-test statistic and p-value matrices.")}
cl5 <- colorRampPalette(c("white", "blue"))
bl5 <- colorRampPalette(c("blue", "red"))

par(mfrow=c(1,2))
tmp <- as.numeric(table(fchannel))
corrplot(ms[ford,ford],is.corr=FALSE,method="color",tl.col=ccol[ford], 
         tl.cex=0.8, mar=c(0,0,1,0),oma=c(0,0,2,0))
title("Test statistic") 
corrRect(tmp,col=Syncol,lwd=4)


corrplot((mp[ford,ford]),is.corr=FALSE,method="color",tl.col=ccol[ford], 
         tl.cex=0.8,col=mycol3[-c(1:2)], mar=c(0,0,1,0),
         p.mat=mp[ford,ford], sig.level=0.05)
title("p-values")
corrRect(tmp,col=Syncol,lwd=4);rm(tmp)
```
The `X`'s in the above right figure denote a p-value greater than 0.05.

## Scatterplots 

We will run PCA on the untransformed correlation matrix so the data can be 
viewed in 2-dimensions.  The colors correspond to synapse type.

```{r cc-pca1,h=7,w=7,fig.cap=fig$cap("pairs1","PCA of untransformed correlation matrix")}
pca1 <- prcomp(corrf,center=TRUE, scale=TRUE)
pairs(pca1$x[,1:3], col=ccol3[ford],pch=20, cex=2)
```


```{r cc_3dCorr, fig.cap=fig$cap("rgl1","PCA 1-3 on untransformed correlation matrix")}
pca <- pca1$x
rgl::plot3d(pca[,1],pca[,2],pca[,3],type='s',col=ccol3[ford], size=1)
rgl::rgl.texts(pca[,1],pca[,2],pca[,3],abbreviate(rownames(pca)), col=ccol3[ford], adj=c(0,2))
subid <- currentSubscene3d()
rglwidget(elementId="rgl-pca1",width=720,height=720)
```


```{r cc-lda1}
d1 <- data.frame(type=exType,pca1$x)
#d1 <- data.frame(type=channel.type2[ford],pca1$x)
try(lda.fit <- lda(type ~ ., data=d1),silent=FALSE)
geterrmessage()
lda.fit <- lda(type ~ ., data=d1[,1:3])
```

<FONT COLOR=#750000> 
LDA was run using only the first 2 out of 24 principal 
components from the untransformed correlation matrix.
</FONT>  


```{r cc-voronoi,h=7,w=8,fig.cap=fig$cap("voronoi","Voronoi diagram on class means from LDA on PCA of untransformed correlation matrix")}

voronoidf <- data.frame(x=lda.fit$means[,1],y=lda.fit$means[,2])

#This creates the voronoi line segments
plot(d1[,2:3], col=ccol3[ford], pch=20, cex=1.5)

text(d1[,2:3], labels=rownames(d1), pos=ifelse(d1[,2]<4,4,2), 
     col=ccol3[ford], cex=1.2)

deldir(x = voronoidf$x,y = voronoidf$y, rw = c(-5,5,-5,5), 
       plotit=TRUE, add=TRUE, wl='te')
text(voronoidf, labels=rownames(voronoidf), pos=1)
```


<footer>
<p> [Back to Top][Data]</p>
</footer>
