---
title: "Synaptome Stats: Marker Exploration"
author: "Jesse Leigh Patsolic"
date: '`r Sys.Date()`'
output:
  html_document:
    fig_caption: yes
    fig_height: 5
    fig_width: 5
    fig_retina: 2.7
    highlight: pygments
    keep_md: yes
    number_sections: yes
    theme: cerulean
    toc: yes
    toc_depth: 3
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
---
```{r render, eval=FALSE, echo=FALSE}
require(rmarkdown)
rm(list=ls()); 
rmarkdown::render(grep("MarkerExploration.Rmd", dir(), value=TRUE))
system('open MarkerExploration.html')
```
```{r setup,include=FALSE,results='asis',message=FALSE}
### Library calls here.
library(rmarkdown)
library(knitr)
require(ggplot2)
require(gridExtra)
require(gplots)
require(lattice)
require(hexbin)
require(data.table)
require(dplyr)
require(energy)
require(lattice)
require(colorRamps)
require(corrplot)
require(doMC); registerDoMC(4)
require(foreach)
```

```{r knitrOpts,include=FALSE,results='asis',message=FALSE,warning=FALSE}
### The following options and figure numbering functions
### were setup by Youngser Park
knitr::opts_chunk$set(cache=FALSE, autodep=TRUE)
dep_auto() # figure out dependencies automatically
opts_chunk$set(cache=FALSE,echo=TRUE,warning=FALSE,message=FALSE,comment="#",fig.path='../Figures/MarkerExploration_figure/', dpi=100,dev=c('png','pdf'))

opts_knit$set(aliases=c(h='fig.height', w='fig.width', cap='fig.cap', scap='fig.scap'))                                                                               
opts_knit$set(eval.after = c('fig.cap','fig.scap'))                                                                            
knit_hooks$set(document = function(x) {
  gsub('(\\\\end\\{knitrout\\}[\n]+)', '\\1\\\\noindent ', x)                                                   })
#opts_knit$set(animation.fun = hook_scianimator)

knit_hooks$set(plot = function(x, options) {
       paste('<figure><img src="',
             opts_knit$get('base.url'), paste(x, collapse = '.'),
             '"><figcaption>', options$fig.cap, '</figcaption></figure>',
             sep = '')
 })
 
 fn = local({
   i = 0
   function(x) {
     i <<- i + 1
     paste('Figure ', i, ': ', x, sep = '')
   }
 })

 fig <- local({
    i <- 0
    ref <- list()
    list(
        cap=function(refName, text) {
            i <<- i + 1
            ref[[refName]] <<- i
            paste("<b>Figure ", i, ": ", text, "</b><br><br>", sep="")
        },
        ref=function(refName) {
            ref[[refName]]
        })
})
```

# Data 


Here we read in the data and select a random half of it for exploration. 

```{r cc-data, eval=TRUE}
featFull <- fread("../data/synapsinR_7thA.tif.Pivots.txt.2011Features.txt",showProgress=FALSE)

### Setting a seed and creating an index vector
### to select half of the data
set.seed(2^10)
half1 <- sample(dim(featFull)[1],dim(featFull)[1]/2)
half2 <- setdiff(1:dim(featFull)[1],half1)

feat <- featFull[half1,]
dim(feat)

## Setting the channel names
channel <- c('Synap_1','Synap_2','VGlut1_t1','VGlut1_t2','VGlut2','Vglut3',
              'psd','glur2','nmdar1','nr2b','gad','VGAT',
              'PV','Gephyr','GABAR1','GABABR','CR1','5HT1A',
              'NOS','TH','VACht','Synapo','tubuli','DAPI')

## Setting the channel types
channel.type <- c('ex.pre','ex.pre','ex.pre','ex.pre','ex.pre','in.pre.small',
                  'ex.post','ex.post','ex.post','ex.post','in.pre','in.pre',
                  'in.pre','in.post','in.post','in.post','in.pre.small','other',
                  'ex.post','other','other','ex.post','none','none')

nchannel <- length(channel)
nfeat <- ncol(feat) / nchannel

## Createing factor variables for channel and channel type sorted properly
ffchannel <- (factor(channel.type,
    levels= c("ex.pre","ex.post","in.pre","in.post","in.pre.small","other","none")
    ))
fchannel <- as.numeric(factor(channel.type,
    levels= c("ex.pre","ex.post","in.pre","in.post","in.pre.small","other","none")
    ))
ford <- order(fchannel)


## Setting up colors for channel types
Syncol <- c("#197300","#5ed155","#660000","#cc0000","#ff9933","mediumblue","gold")
ccol <- Syncol[fchannel]

exType <- factor(c(rep("ex",11),rep("in",6),rep("other",7)),ordered=TRUE)
exCol<-exType;levels(exCol) <- c("#197300","#990000","mediumblue");
exCol <- as.character(exCol)

fname <- as.vector(sapply(channel,function(x) paste0(x,paste0("F",0:5))))
names(feat) <- fname
fcol <- rep(ccol, each=6)
mycol <- colorpanel(100, "purple", "black", "green")
mycol2 <- matlab.like(nchannel)
```

## Data transformations

```{r cc-datatrans, eval=TRUE}
f <- lapply(1:6,function(x){seq(x,ncol(feat),by=nfeat)})
featF <- lapply(f,function(x){subset(feat,select=x)})

featF0 <- featF[[1]]
f01e3 <- 1e3*data.table(apply(X=featF0,2,function(x){((x-min(x))/(max(x)-min(x)))}))

fs <- f01e3

### Taking log_10 on data + 1.
log1f <- log10(featF0 + 1)
slog1f <- data.table(scale(log1f, center=TRUE,scale=TRUE))
```

We now have the following data sets:

- `featF0`: The feature vector looking only at the integrated brightness features.
- `fs`:  The feature vector scaled between $[0,1000]$.
- `logf1`: The feature vector, plus one, then $log_{10}$ is applied. 
- `slog1f`: The feature vector, plus one, $log_{10}$, then scaled by
  subtracting the mean and dividing by the sample standard deviation.

# Marker Exploration 


## Top 1% of Synap_1

```{r cc-top1synap1F0}
top1synap1F0 <- slog1f[Synap_1F0 > quantile(Synap_1F0, 0.99),]
dim(top1synap1F0)

lp1 <- splom(~top1synap1F0,
	panel=panel.hexbinplot, colramp=BTC, 
	diag.panel = function(x, ...){
	 yrng <- current.panel.limits()$ylim
	 d <- density(x, na.rm=TRUE)
	 d$y <- with(d, yrng[1] + 0.95 * diff(yrng) * y / max(y) )
	 panel.lines(d)
	 diag.panel.splom(x, ...)
	 },
	lower.panel = function(x, y, ...){
	 panel.hexbinplot(x, y, ...)
	 panel.loess(x, y, ..., col = 'red')
	 },
	pscale=0,varname.cex=0.9
 )

pdf("../Figures/MarkerExploration_figure/cc-top1synap1F0.pdf",height=25,width=25)
print(lp1)
dev.off()
```

Note that when we select only the top 1% of Synap_1F0 levels we end up with only `r dim(top1synap1F0)[1]` 
observations.  
This plot is too large to show inline, you can view it here:
[F0 Lattice Plot conditioned on top 1% of Synap_1F0](../Figures/MarkerExploration_figure/cc-top1synap1F0.pdf)


```{r cc-top1gadF0}
top1gadF0 <- slog1f[gadF0 > quantile(gadF0, 0.99),]
dim(top1gadF0)

lp2 <- splom(~top1gadF0,
	panel=panel.hexbinplot, colramp=BTC, 
	diag.panel = function(x, ...){
	 yrng <- current.panel.limits()$ylim
	 d <- density(x, na.rm=TRUE)
	 d$y <- with(d, yrng[1] + 0.95 * diff(yrng) * y / max(y) )
	 panel.lines(d)
	 diag.panel.splom(x, ...)
	 },
	lower.panel = function(x, y, ...){
	 panel.hexbinplot(x, y, ...)
	 panel.loess(x, y, ..., col = 'red')
	 },
	pscale=0,varname.cex=0.9
 )

pdf("../Figures/MarkerExploration_figure/cc-top1gadF0.pdf",height=25,width=25)
print(lp2)
dev.off()
```

This plot is too large to show inline, you can view it here:
[F0 Lattice Plot conditioned on top 1% of gadF0](../Figures/MarkerExploration_figure/cc-top1gadF0.pdf)

## Correlation Matrix of markers 

```{r cc_corRawF0,w=7,h=7, eval=TRUE,fig.cap=fig$cap("corr","Correlation on untransformed F0 data, reordered by synapse type.")}
tmp <- as.numeric(table(fchannel))
corrf <- cor(featF0)[ford,ford]
corrplot(corrf,method="color",tl.col=ccol[ford], tl.cex=0.8)
```

```{r cc_corRawF0-5,w=24,h=24, eval=TRUE,fig.cap=fig$cap("corr","Correlation on untransformed data F0-5, reordered by synapse type.")}
tmp <- as.numeric(table(fchannel))
bford <- order(rep(fchannel,each=6))
nord <- Reduce('c', f)
cr <- rep(ccol, each=6)
corrf <- cor(feat)[bford,bford]
#corrf <- cor(feat)[nord,nord]
corrplot(corrf,method="color",tl.col=cr[bford],tl.cex=0.8)
#corrplot(corrf,method="color",tl.col=cr[nord],tl.cex=0.8)
```


## Distance Covariance Test

```{r cc-dcov,eval=FALSE,echo=FALSE}
set.seed(317)
sam1 <- sample(dim(featF0)[1], 1e3)
tmp <- as.data.frame((featF0[sam1,]))

set.seed(331)

combcols <- t(combn(24,2))

    
dc <- foreach(i = 1:dim(combcols)[1]) %dopar% {
      dcov.test(x=tmp[,combcols[i,1]],y=tmp[,combcols[i,2]])
        }

ms <- matrix(as.numeric(0),24,24)
mp <- matrix(as.numeric(0),24,24)

for(i in 1:length(dc)){
    ms[combcols[i,1],combcols[i,2]] <- dc[[i]]$statistic
    ms[combcols[i,2],combcols[i,1]] <- dc[[i]]$statistic
    mp[combcols[i,1],combcols[i,2]] <- dc[[i]]$p
    mp[combcols[i,2],combcols[i,1]] <- dc[[i]]$p
}

rownames(ms) <- colnames(featF0)
rownames(mp) <- colnames(featF0)
colnames(ms) <- colnames(featF0)
colnames(mp) <- colnames(featF0)

diag(ms) <- as.numeric(10)
diag(mp) <- as.numeric(1)


tmp <- as.numeric(table(fchannel))
par(mfrow=c(1,2))
corrplot(log10(mp)[ford,ford],is.corr=FALSE,method="color",tl.col=ccol[ford], tl.cex=0.8)
corrRect(tmp,col=Syncol,lwd=4)
corrplot(ms[ford,ford],is.corr=FALSE,method="color",tl.col=ccol[ford], tl.cex=0.8)
corrRect(tmp,col=Syncol,lwd=4)
```


<footer>
<p> [Back to Top][Outline]</p>
</footer>

