---
title: "Synapse Clustering: Random Half"
author: "JLP"
date: '`r Sys.Date()`'
output:
  html_document:
    fig_caption: yes
    fig_height: 5
    fig_width: 5
    highlight: pygments
    keep_md: yes
    number_sections: yes
    theme: cerulean
    toc: yes
    toc_depth: 3
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
---
```{r render, eval=FALSE, echo=FALSE}
require(rmarkdown)
rm(list=ls()); 
### SAVE DOCUMENT!!!
rmarkdown::render("./randomHalf.Rmd");
system('open randomHalf.html')
```
```{r setup,include=FALSE,results='asis'}
suppressMessages(require(Matrix))
suppressMessages(require(colorRamps))
suppressMessages(require(corrplot))
suppressMessages(require(gplots))
suppressMessages(require(dplyr))
suppressMessages(require(reshape2))
suppressMessages(require(ggplot2))
suppressMessages(require(parallel))
suppressMessages(require(data.table))
suppressMessages(require(fpc))
suppressMessages(require(repr))
suppressMessages(require(doMC))
suppressMessages(require(mclust))
suppressMessages(require(rgl))
suppressMessages(require(rglwidget))
source("./bic.r")
source("./clusterFraction.r")
source("http://www.cis.jhu.edu/~parky/Synapse/getElbows.R")
registerDoMC(6)

library(rmarkdown)
library(knitr)

knitr::opts_chunk$set(cache=TRUE, autodep=TRUE)
dep_auto() # figure out dependencies automatically
opts_chunk$set(cache=FALSE,echo=TRUE,warning=FALSE,message=FALSE,comment="#",fig.path='figure/',
               dpi=100,dev=c('png','pdf'))

opts_knit$set(aliases=c(h='fig.height', w='fig.width', cap='fig.cap', scap='fig.scap'))                                                                               
opts_knit$set(eval.after = c('fig.cap','fig.scap'))                                                                            
knit_hooks$set(document = function(x) {
  gsub('(\\\\end\\{knitrout\\}[\n]+)', '\\1\\\\noindent ', x)                                                   })
#opts_knit$set(animation.fun = hook_scianimator)

knit_hooks$set(webgl=hook_webgl)
knit_hooks$set(plot = function(x, options) {
       paste('<figure><img src="',
             opts_knit$get('base.url'), paste(x, collapse = '.'),
             '"><figcaption>', options$fig.cap, '</figcaption></figure>',
             sep = '')
 })
 
 fn = local({
   i = 0
   function(x) {
     i <<- i + 1
     paste('Figure ', i, ': ', x, sep = '')
   }
 })

 fig <- local({
    i <- 0
    ref <- list()
    list(
        cap=function(refName, text) {
            i <<- i + 1
            ref[[refName]] <<- i
            paste("<b>Figure ", i, ": ", text, "</b><br><br>", sep="")
        },
        ref=function(refName) {
            ref[[refName]]
        })
})

```

Previous work by Youngser Park can be found
[here.](http://www.cis.jhu.edu/~parky/Synapse/synapse.html)

# Introduction

# Data
> The corresponds to 24 channels x 6 features per synapse, ordered like
> c0f0,c0f1,c0f2,c0f3,c0f4,c0f5,c1f0,c1f1... etc
>
>f0 = integrated brightness  
>f1 = local brightness  
>f2 = distance to Center of Mass  
>f3 = moment of inertia around synapsin maxima  
>f4,f5 are features that I forget what they are.. would need to ask brad.   
>i would throw them out, I did so in my kohonen code (which you have, its in matlab).

and

> On Feb 8, 2016, at 2:00 PM, Kristina Micheva <kmicheva@stanford.edu> wrote:

* <FONT COLOR=#197300>_Excitatory presynaptic: 'Synap', 'Synap', 'VGlut1', 'VGlut1', 'VGlut2'_</FONT>,
* <FONT COLOR=#5ed155>_Excitatory postsynaptic: 'psd', 'glur2', 'nmdar1', 'nr2b', 'NOS', 'Synapo'_</FONT> (but further away than PSD, gluR2, nmdar1 and nr2b)
* <FONT COLOR=#990000>_Inhibitory presynaptic: 'gad', 'VGAT', 'PV'_</FONT>,
* <FONT COLOR=#ff3333>_Inhibitory postsynaptic: 'Gephyr', 'GABAR1', 'GABABR', 'NOS'_</FONT>,
* <FONT COLOR=#ff9933>_At a very small number of inhibitory: 'Vglut3' (presynaptic), 'CR1'(presynaptic)_</FONT>,
* <FONT COLOR="mediumblue">_Other synapses:'5HT1A', 'TH', 'VACht'_</FONT>,
* <FONT COLOR="gold">_Not at synapses: 'tubuli', 'DAPI'_</FONT>.

and 

> On March 10, 2016, 00:29:04 (UTC), Kristina Micheva <kmicheva@stanford.edu> wrote:

There are 2 different Synap channels (2 different antibodies were
used), so that part is fine.
And 2 different VGluT1 channels (same antibody but done at different
times)
The NOS channel is the same, so count it as one even though it appears
twice. It is listed two times because it can be found at both excitatory
and inhibitory synapses. This is where your count of 25 comes, even
though there are 24 channels.
I would also add the 2 Synap channels to the Inhibitory presynaptic
category - there is supposed to be synapsin there, but at lower levels
compared to excitatory presynaptic category.

- Note:  The order of the channels are given by line `227` in the `kohenen.m` file which can be found in the dropbox folder. 
- `Synap` and `Synap` have been augmented to `Synap_1` and `Synap_2` for clarity. 
- `VGlut1` and `VGlut1` have been augmented to `VGlut1_t1` and `VGlut1_t2` to distinguish between the different times (which are unknown).

```{r cc_data, eval=TRUE}
feat <- fread("../Data/synapsinR_7thA.tif.Pivots.txt.2011Features.txt",showProgress=FALSE)
dim(feat)

### Setting a seed and creating an index vector
### to select half of the data
set.seed(2^13)
half1 <- sample(dim(feat)[1],dim(feat)[1]/2)

feat <- feat[half1,]
channel <- c('Synap_1','Synap_2','VGlut1_t1','VGlut1_t2','VGlut2','Vglut3',
              'psd','glur2','nmdar1','nr2b','gad','VGAT',
              'PV','Gephyr','GABAR1','GABABR','CR1','5HT1A',
              'NOS','TH','VACht','Synapo','tubuli','DAPI')

## NOS as ex.post
#channel.type <- c('ex.pre','ex.pre','ex.pre','ex.pre','ex.pre','in.pre.small',
#                  'ex.post','ex.post','ex.post','ex.post','in.pre','in.pre',
#                  'in.pre','in.post','in.post','in.post','in.pre.small','other',
#                  'ex.post','other','other','ex.post','none','none')
#
## NOS as in.post
channel.type <- c('ex.pre','ex.pre','ex.pre','ex.pre','ex.pre','in.pre.small',
                  'ex.post','ex.post','ex.post','ex.post','in.pre','in.pre',
                  'in.pre','in.post','in.post','in.post','in.pre.small','other',
                  'in.post','other','other','ex.post','none','none')
nchannel <- length(channel)
nfeat <- ncol(feat) / nchannel
fchannel <- as.numeric(factor(channel.type,
    levels= c("ex.pre","ex.post","in.pre","in.post","in.pre.small","other","none")
    ))
ford <- order(fchannel)
#ccol <- rainbow(max(fchannel))[fchannel]
Syncol <- c("#197300","#5ed155","#990000","#cc0000","#ff9933","mediumblue","gold")
ccol <- Syncol[fchannel]

exType <- factor(c(rep("ex",11),rep("in",6),rep("other",7)),ordered=TRUE)
exCol<-exType;levels(exCol) <- c("#197300","#990000","mediumblue");
exCol <- as.character(exCol)

fname <- as.vector(sapply(channel,function(x) paste0(x,paste0("F",0:5))))
names(feat) <- fname
fcol <- rep(ccol, each=6)
mycol <- colorpanel(100, "purple", "black", "green")
mycol2 <- matlab.like(nchannel)
```

## Transformations

Each of the channels has an arbitrary independent linear transformation and thus
converting to z-scores or quantiles is required.


Selecting only the `f0` features, we will use both the raw data and a
transformed version, transforming by adding 1 and taking
the log base 10 for each column and scaling. 

```{r cc_featF0, eval=TRUE}
f0 <- seq(1,ncol(feat),by=nfeat)
fRaw <- subset(feat, select=f0)
#fLog <- apply(X=(fRaw+1),2,log10)
#fLogScaled <- scale(flog, center=TRUE, scale=TRUE)

eps <- 2*.Machine$double.eps
f01 <- apply(X=fRaw, 2, function(x){((x-min(x))/(max(x)-min(x)))})
fRank <- apply(fRaw,2,rank,ties.method='average')
fLog <- apply(f01,2,function(x){ log10(x+eps)})
```

# Correlations 

## Raw

```{r cc_corRaw, eval=TRUE, fig.cap=fig$cap("corRaw", "Correlation on raw data, reordered by synapse type.")}
tmp <- as.numeric(table(fchannel))
cmatRaw <- cor(fRaw)
corrplot(cmatRaw[ford,ford],method="color",tl.col=ccol[ford])
corrRect(tmp,col=Syncol,lwd=4)
```

## Log

```{r cc_corLog, eval=TRUE, fig.cap=fig$cap("corLog", "Correlation on log data, reordered by synapse type.")}
cmatLog <- cor(fLog)
corrplot(cmatLog[ford,ford],method="color",tl.col=ccol[ford])
corrRect(tmp,col=Syncol,lwd=4)
```

## Rank

```{r cc_corRank, eval=TRUE, fig.cap=fig$cap("corRank", "Correlation on rank data, reordered by synapse type.")}
cmatRank <- cor(fRank)
corrplot(cmatRank[ford,ford],method="color",tl.col=ccol[ford])
corrRect(tmp,col=Syncol,lwd=4)
```

# PCA and Scree Plots on the Correlation Matrices

Note that centering and scaling are set to `FALSE`.

```{r pcaRaw, eval=TRUE}
pcaRaw <- prcomp(cmatRaw,scale=FALSE, center=FALSE)
elRaw <- getElbows(pcaRaw$sdev, plot=FALSE) 
```
```{r pcaLog, eval=TRUE}
pcaLog <- prcomp(cmatLog,scale=FALSE, center=FALSE)
elLog <- getElbows(pcaLog$sdev, plot=FALSE) 
```
```{r pcaRank, eval=TRUE}
pcaRank <- prcomp(cmatRank,scale=FALSE, center=FALSE)
elRank <- getElbows(pcaRank$sdev, plot=FALSE)
```

```{r cc-scree-plots, eval=TRUE, w=8,h=4, fig.cap=fig$cap("scree-plots","Scree plots")}
par(mfrow=c(1,3))
plot(pcaRaw$sdev, type='b'); title("Scree plot: Raw Cor.  Matrix")
points(cbind(elRaw,pcaRaw$sdev[elRaw]), col='red', pch=20)
plot(pcaLog$sdev, type='b'); title("Scree plot: Log Cor. Matrix")
points(cbind(elLog,pcaLog$sdev[elLog]), col='red', pch=20)
plot(pcaRank$sdev, type='b'); title("Scree plot: Rank Cor. Matrix")
points(cbind(elRank,pcaRank$sdev[elRank]), col='red', pch=20)
```



# PAMK

Running `pamk` for $K = 7$.

## PAMK on Raw data
```{r cc_pamkRaw, eval=TRUE, fig.cap=fig$cap("silRaw", "Average silhouette width plot on Raw Cor. Matrix")}
K <- 7
pamRaw7 <- pamk(pcaRaw$x[,1:elRaw[2]],krange=K,usepam=FALSE,critout=FALSE)
```


```{r cc_aggRaw, eval=TRUE}
aggRaw <- aggregate(pcaRaw$x[,1:elRaw[2]],by=list(lab=pamRaw7$pamo$clustering),FUN=mean)

aggRaw <- as.matrix(aggRaw[,-1])
rownames(aggRaw) <- clusterFraction(pamRaw7$pamo$clustering)
```

```{r cc_heatmapRaw,eval=TRUE,w=6,h=6,echo=TRUE,fig.cap=fig$cap("heat-raw", "Heat Map of Raw Data")}
heatmap.2(as.matrix(aggRaw), trace="none",col=mycol,cexRow=0.8, keysize=1,symkey=FALSE,symbreaks=FALSE,scale="none", srtCol=90) # 
```

## PAMK on Log Data

```{r cc_pamkLog, eval=TRUE, fig.cap=fig$cap("silLog", "Average silhouette width plot on Log Cor. Matrix")}
pamLog7 <- pamk(pcaLog$x[,1:elLog[2]],krange=K,usepam=FALSE,critout=FALSE)
#plot(pamLog5$crit, type='b')
```

```{r cc_aggLog, eval=TRUE}
aggLog <- aggregate(pcaLog$x[,1:elLog[2]],by=list(lab=pamLog7$pamo$clustering),FUN=mean)

aggLog <- as.matrix(aggLog[,-1])
rownames(aggLog) <- clusterFraction(pamLog7$pamo$clustering)
```

```{r cc_heatmapLog,eval=TRUE,w=6,h=6,echo=TRUE,fig.cap=fig$cap("heat-log", "Heat Map of Log Data")}
heatmap.2(as.matrix(aggLog), trace="none",col=mycol,cexRow=0.8, keysize=1,symkey=FALSE,symbreaks=FALSE,scale="none", srtCol=90) # 
```

## PAMK on Ranked Data

```{r cc_pamkRank, eval=TRUE, fig.cap=fig$cap("silRank", "Average silhouette width plot on Rank Cor. Matrix")}
pamRank7 <- pamk(pcaRank$x[,1:elRank[2]],krange=K,usepam=FALSE,critout=FALSE)
#plot(pamRank5$crit, type='b')
```

```{r cc_aggRank, eval=TRUE}
aggRank <- aggregate(pcaRank$x[,1:elRank[2]],by=list(lab=pamRank7$pamo$clustering),FUN=mean)

aggRank <- as.matrix(aggRank[,-1])
rownames(aggRank) <- clusterFraction(pamRank7$pamo$clustering)
```

```{r cc_heatmapRank,eval=TRUE,w=6,h=6,echo=TRUE,fig.cap=fig$cap("heat-rank", "Heat Map of Ranked Data")}
heatmap.2(as.matrix(aggRank), trace="none",col=mycol,cexRow=0.8, keysize=1,symkey=FALSE,symbreaks=FALSE,scale="none", srtCol=90) # 
```


## ARI

We are considering the channel types to be the "ground truth" here.

```{r cc_truth}
truth <- as.numeric(factor(channel.type))
```

Calculating the ARI for the raw data.

```{r ariRaw}
xRaw <- as.numeric(pamRaw7$pamo$clustering)
ariRaw <- foreach(i = 1:1e4, .combine='c') %dopar%{
    y <- sample(xRaw)
    mclust::adjustedRandIndex(truth,y)
    }
```


```{r ariLog}
xLog <- as.numeric(pamLog7$pamo$clustering)
ariLog <- foreach(i = 1:1e4, .combine='c') %dopar%{
    y <- sample(xLog)
    mclust::adjustedRandIndex(truth,y)
    }
```

```{r ariRank}
xRank <- as.numeric(pamRank7$pamo$clustering)
ariRank <- foreach(i = 1:1e4, .combine='c') %dopar%{
    y <- sample(xRank)
    mclust::adjustedRandIndex(truth,y)
    }
```

```{r cc_ariPlots, w=9,h=4, fig.cap=fig$cap("ariPlot","ARI plots")}
par(mfrow=c(1,3))
hist(ariRaw)
abline(v = adjustedRandIndex(truth,xRaw), lwd=2, col='red')
hist(ariLog)
abline(v = adjustedRandIndex(truth,xLog), lwd=2, col='red')
hist(ariRank)
abline(v = adjustedRandIndex(truth,xRank), lwd=2, col='red')
```

```{r cc_ariNum}
adjustedRandIndex(truth, xRaw)
adjustedRandIndex(truth, xLog)
adjustedRandIndex(truth, xRank)
```


## ARI vs Embedding dimension.

Here we investigeta how ARI varies with the embedding dimension.

### Functions
```{r cc_f1}
avd <- function(pcaObj,elbow,truth){
###
### Given a pca object with an elbow as upperbound
### compute the ARI of truth versus the pamk clustering for each
### dimension from 1:elbow
###
require(foreach)

out <- foreach(i = 1:elbow, .combine="rbind") %do% {
    pmk <- pamk(pcaObj$x[,1:i], krange=K, usepam=FALSE, critout=FALSE)
    data.frame(Embedding_Dimension=i,ARI=adjustedRandIndex(truth, pmk$pamo$clustering))
    }
return(out)
}
```

### ARI Plots

```{r arivdimRaw, eval=TRUE, w=9,h=4, fig.cap=fig$cap("ariVdim", "ARI vs DIM")}
avdRaw <- avd(pcaRaw, elRaw[2], truth)
avdLog <- avd(pcaLog, elLog[2], truth)
avdRank <- avd(pcaRank, elRank[2], truth)

avdRaw$group <- "Raw"
avdLog$group <- "Log"
avdRank$group <- "Rank"

avdDat <- data.frame(rbind(avdRaw, avdLog, avdRank))
#par(mfrow=c(1,3))
#plot(avdRaw, type='b', main="Raw Data")
#plot(avdLog, type='b', main="Log Data")
#plot(avdRank, type='b', main="Rank Data")

p1 <- ggplot(data=avdDat, aes(x=Embedding_Dimension, y=ARI,group=group,color=group)) + 
        geom_line(size=1.5) + geom_point(size=3) +
            theme(axis.text=element_text(size=18)) + ggtitle("ARI vs. DIM")
print(p1)
```

# Plots of embeddings

First let's look at the 3-d embeddings and make some decisions from
there.

```{r cc_3dEmb, h=6,w=6}
pairs(pcaRaw$x[,1:3],col=exCol,pch=as.numeric(exType)+15,cex=2,main="Raw")
pairs(pcaLog$x[,1:3],col=exCol, pch=as.numeric(exType)+15,cex=2,main="Log")
pairs(pcaRank$x[,1:3],col=exCol,pch=as.numeric(exType)+15,cex=2,main="Rank")
```

```{r cc_3dRaw}
pca <- pcaRaw$x[,1:3]
rgl::plot3d(pca[,1],pca[,2],pca[,3],type='s', col=exCol, size=1, main="Raw")
rgl::rgl.texts(pca[,1],pca[,2],pca[,3], channel[ford], col=exCol, adj=c(0,1.5))
subid <- currentSubscene3d()
rglwidget(elementId="plot3dRaw")
```

```{r cc_3dLog}
pca <- pcaLog$x
rgl::plot3d(pca[,1],pca[,2],pca[,3],type='s', col=exCol, size=1, main="Log")
rgl::rgl.texts(pca[,1],pca[,2],pca[,3], channel[ford], col=exCol, adj=c(0,1.5))
subid <- currentSubscene3d()
rglwidget(elementId="plot3dLog")
```

```{r cc_3dRank}
pca <- pcaRank$x[,1:3]
rgl::plot3d(pca[,1],pca[,2],pca[,3],type='s', col=exCol, size=1, main="Rank")
rgl::rgl.texts(pca[,1],pca[,2],pca[,3], channel[ford], col=exCol, adj=c(0,1.5))
subid <- currentSubscene3d()
rglwidget(elementId="plot3dRank")
```


# Pick $K=3$.

## PAMK=3 on Raw data
```{r cc_pamk3Raw, eval=TRUE, fig.cap=fig$cap("silRaw", "Average silhouette width plot on Raw Cor. Matrix")}
K <- 3
pamRaw3 <- pamk(pcaRaw$x[,1:elRaw[2]],krange=K,usepam=FALSE,critout=FALSE)
```

```{r cc_agg3Raw, eval=TRUE}
agg3Raw <- aggregate(pcaRaw$x[,1:elRaw[2]],by=list(lab=pamRaw3$pamo$clustering),FUN=mean)

agg3Raw <- as.matrix(agg3Raw[,-1])
rownames(agg3Raw) <- clusterFraction(pamRaw3$pamo$clustering)
```

```{r cc_heatmap3Raw,eval=TRUE,w=6,h=6,echo=TRUE,fig.cap=fig$cap("heat-3raw", "Heat Map of Raw Data")}
heatmap.2(as.matrix(agg3Raw), trace="none",col=mycol,cexRow=0.8, keysize=1,symkey=FALSE,symbreaks=FALSE,scale="none", srtCol=90) # 
```

## PAMK=3 on Log Data

```{r cc_pamk3Log, eval=TRUE, fig.cap=fig$cap("silLog", "Average silhouette width plot on Log Cor. Matrix")}
pamLog3 <- pamk(pcaLog$x[,1:elLog[2]],krange=K,usepam=FALSE,critout=FALSE)
#plot(pamLog5$crit, type='b')
```

```{r cc_agg3Log, eval=TRUE}
agg3Log <- aggregate(pcaLog$x[,1:elLog[2]],by=list(lab=pamLog3$pamo$clustering),FUN=mean)

agg3Log <- as.matrix(agg3Log[,-1])
rownames(agg3Log) <- clusterFraction(pamLog3$pamo$clustering)
```

```{r cc_heatmap3Log,eval=TRUE,w=6,h=6,echo=TRUE,fig.cap=fig$cap("heat-3log", "Heat Map of Log Data")}
heatmap.2(as.matrix(agg3Log), trace="none",col=mycol,cexRow=0.8, keysize=1,symkey=FALSE,symbreaks=FALSE,scale="none", srtCol=90) # 
```

## PAMK=3 on Ranked Data

```{r cc_pamk3Rank, eval=TRUE, fig.cap=fig$cap("silRank", "Average silhouette width plot on Rank Cor. Matrix")}
pamRank3 <- pamk(pcaRank$x[,1:elRank[2]],krange=K,usepam=FALSE,critout=FALSE)
#plot(pamRank5$crit, type='b')
```

```{r cc_agg3Rank, eval=TRUE}
agg3Rank <- aggregate(pcaRank$x[,1:elRank[2]],by=list(lab=pamRank3$pamo$clustering),FUN=mean)

agg3Rank <- as.matrix(agg3Rank[,-1])
rownames(agg3Rank) <- clusterFraction(pamRank3$pamo$clustering)
```

```{r cc_heatmap3Rank,eval=TRUE,w=6,h=6,echo=TRUE,fig.cap=fig$cap("heat-3rank", "Heat Map of Ranked Data")}
heatmap.2(as.matrix(agg3Rank), trace="none",col=mycol,cexRow=0.8, keysize=1,symkey=FALSE,symbreaks=FALSE,scale="none", srtCol=90) # 
```

## ARI Plots for $K=3$

```{r arivdim3Raw, eval=TRUE, w=9,h=4, fig.cap=fig$cap("ariVdim", "ARI vs DIM")}
avd3Raw <- avd(pcaRaw,elRaw[2],truth)
avd3Log <- avd(pcaLog,elLog[2],truth)
avd3Rank <- avd(pcaRank,elRank[2],truth)

avd3Raw$group <- "Raw"
avd3Log$group <- "Log"
avd3Rank$group <- "Rank"

avd3Dat <- data.frame(rbind(avd3Raw, avd3Log, avd3Rank))

p2 <- ggplot(data=avd3Dat, aes(x=Embedding_Dimension, y=ARI,group=group,color=group)) + 
        geom_line(size=1.5) + geom_point(size=3) +
            theme(axis.text=element_text(size=18)) + ggtitle("ARI vs. DIM")
print(p2)
```

