---
title: "Synapse Stats: Synapse Exploration"
author: "JLP"
date: '`r Sys.Date()`'
output:
  html_document:
    fig_caption: yes
    fig_height: 5
    fig_width: 5
    highlight: pygments
    keep_md: yes
    number_sections: yes
    theme: cerulean
    toc: yes
    toc_depth: 3
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
---
```{r render, eval=FALSE, echo=FALSE}
require(rmarkdown)
rm(list=ls()); 
rmarkdown::render(grep("SynapseExploration.Rmd", dir(), value=TRUE))
system('open SynapseExploration.html')
```
```{r setup,include=FALSE,results='asis',message=FALSE,warning=FALSE}
### Library calls here.
require(Matrix)
require(colorRamps)
require(RColorBrewer)
require(hexbin)
require(corrplot)
require(gplots)
require(dplyr)
require(reshape2)
require(ggplot2)
require(gridExtra)
require(lattice)
require(parallel)
require(data.table)
require(fpc)
require(repr)
require(mclust)
require(doMC)
require(rgl)
require(rglwidget)

source("./bic.r")
source("./clusterFraction.r")
source("./kmpp.r")
#source("./getElbows.r")
source("http://www.cis.jhu.edu/~parky/Synapse/getElbows.R")
registerDoMC(6)

library(rmarkdown)
library(knitr)
```
```{r knitr-setup, include=FALSE, results='asis'}
### The following options and figure numbering functions
### were setup by Youngser Park
knitr::opts_chunk$set(cache=TRUE, autodep=TRUE)
dep_auto() # figure out dependencies automatically
opts_chunk$set(cache=FALSE, echo=TRUE, 
               warning=FALSE, message=FALSE, 
               comment="#", fig.path='../Figures/SynapseExploration_figure/',
               dpi=227,dev=c('png','pdf'))

opts_knit$set(aliases = c(h = 'fig.height', w = 'fig.width', 
                          cap='fig.cap', scap='figscap'))

opts_knit$set(eval.after = c('fig.cap','fig.scap'))

knit_hooks$set(document = function(x) {
  gsub('(\\\\end\\{knitrout\\}[\n]+)', '\\1\\\\noindent ', x)
  }
)
#opts_knit$set(animation.fun = hook_scianimator)

knit_hooks$set(plot = function(x, options) {
  paste('<figure><img src="',
        opts_knit$get('base.url'), 
        paste(x, collapse = '.'),
             '"><figcaption>', options$fig.cap, '</figcaption></figure>',
             sep = '')
 })
 
 fn = local({
   i = 0
   function(x) {
     i <<- i + 1
     paste('Figure ', i, ': ', x, sep = '')
   }
 })

 fig <- local({
    i <- 0
    ref <- list()
    list(
        cap=function(refName, text) {
            i <<- i + 1
            ref[[refName]] <<- i
            paste("<b>Figure ", i, ": ", text, "</b><br><br>", sep="")
        },
        ref=function(refName) {
            ref[[refName]]
        })
})
```

[Homepage](http://docs.neurodata.io/synaptome-stats/)  
The formatted source code for this file is 
[here](https://github.com/neurodata/synaptome-stats/blob/gh-pages/Code/SynapseExploration.Rmd).  
And a [raw version here](https://raw.githubusercontent.com/neurodata/synaptome-stats/gh-pages/Code/SynapseExploration.Rmd).    
Previous work by Youngser Park can be found [here](http://www.cis.jhu.edu/~parky/Synapse/synapse.html).  


# Introduction

Following from previous pages, this page will focus on filtering the data 
before clustering to explore if filtering improves the outcome of clustering.

# Data 
Here we read in the data and select a random half of it for exploration. 

```{r cc-data, eval=TRUE}
featFull <- fread("../data/synapsinR_7thA.tif.Pivots.txt.2011Features.txt",
                  showProgress=FALSE)
locFull <- fread("../data/synapsinR_7thA.tif.Pivots.txt",showProgress=FALSE)

### Setting a seed and creating an index vector
### to select half of the data
set.seed(2^10)
half1 <- sample(dim(featFull)[1],dim(featFull)[1]/2)
half2 <- setdiff(1:dim(featFull)[1],half1)

feat <- featFull[half1,]
loc <- locFull[half1,]
dim(feat)

## Setting the channel names
channel <- c('Synap_1','Synap_2','VGlut1_t1','VGlut1_t2','VGlut2','Vglut3',
              'psd','glur2','nmdar1','nr2b','gad','VGAT',
              'PV','Gephyr','GABAR1','GABABR','CR1','5HT1A',
              'NOS','TH','VACht','Synapo','tubuli','DAPI')

## Setting the channel types
channel.type <- c('ex.pre','ex.pre','ex.pre','ex.pre','ex.pre','in.pre.small',
                  'ex.post','ex.post','ex.post','ex.post','in.pre','in.pre',
                  'in.pre','in.post','in.post','in.post','in.pre.small','other',
                  'ex.post','other','other','ex.post','none','none')

nchannel <- length(channel)
nfeat <- ncol(feat) / nchannel

## Createing factor variables for channel and channel type sorted properly
ffchannel <- (factor(channel.type,
    levels= c("ex.pre","ex.post","in.pre","in.post","in.pre.small","other","none")
    ))
fchannel <- as.numeric(factor(channel.type,
    levels= c("ex.pre","ex.post","in.pre","in.post","in.pre.small","other","none")
    ))
ford <- order(fchannel)


## Setting up colors for channel types
Syncol <- c("#197300","#5ed155","#660000","#cc0000","#ff9933","mediumblue","gold")
ccol <- Syncol[fchannel]

exType <- factor(c(rep("ex",11),rep("in",6),rep("other",7)),ordered=TRUE)
exCol<-exType;levels(exCol) <- c("#197300","#990000","mediumblue");
exCol <- as.character(exCol)

fname <- as.vector(sapply(channel,function(x) paste0(x,paste0("F",0:5))))
names(feat) <- fname
fcol <- rep(ccol, each=6)
mycol <- colorpanel(100, "purple", "black", "green")
mycol2 <- matlab.like(nchannel)
```

## Data transformations

```{r cc-datatrans, eval=TRUE}
f <- lapply(1:6,function(x){seq(x,ncol(feat),by=nfeat)})
featF <- lapply(f,function(x){subset(feat,select=x)})

featF0 <- featF[[1]]
f01e3 <- 1e3*data.table(apply(X=featF0, 2, function(x){((x-min(x))/(max(x)-min(x)))}))

fs <- f01e3

### Taking log_10 on data with 0's removed
ans <- apply(featF0, 1, function(row){ any(row == 0)})

logF0 <- log10(featF0[!ans,])
slogF0 <- logF0[,lapply(.SD,scale, center=TRUE,scale=TRUE)]
```

We now have the following data sets:

- `featF0`: The feature vector looking only at the integrated brightness features.
- `fs`:  The feature vector scaled between $[0,1000]$.
- `logF0`: The feature vector, with 0's removed, then $log_{10}$ is applied. 
- `slogF0`: The feature vector, with 0's removed, then $log_{10}$, then scaled by
  subtracting the mean and dividing by the sample standard deviation.



# Kernel Density Estimates of the marginals

```{r synEx-kde1, eval=TRUE,echo=TRUE,cache=FALSE,w=18,h=12,fig.cap=fig$cap("synEx-kde1","Kernel density estimates for each channel, on `fs` data.")}
df1 <- melt(as.matrix(fs))
names(df1) <- c("ind","channel","value")
df1$type <- factor(rep(ffchannel,each=dim(fs)[1]),levels=levels(ffchannel))

lvo <- c(1:5,7:10,19,22,11:16,6,17,18,20,21,23,24)
levels(df1$channel)<-levels(df1$channel)[lvo]

ts <- 22

gg1 <- ggplot(df1, aes(x=value)) + 
    scale_color_manual(values=ccol[lvo]) +
    scale_fill_manual(values=ccol[lvo]) +
    geom_histogram(aes(y=..density..,group=channel,colour=channel),bins=100) +
    geom_density(aes(group=channel, color=channel),size=1.5) +
    facet_wrap( ~ channel, scale='free', ncol=6) +
    theme(plot.title=element_text(size=ts),
          axis.title.x=element_text(size=ts),
          axis.title.y=element_text(size=ts),
          legend.title=element_text(size=ts),
          legend.text=element_text(size=ts-2),
          axis.text=element_text(size=ts-2),
          strip.text=element_text(size=ts), 
          legend.position='none')+
    ggtitle("Kernel Density Estimates of `fs` data.")

print(gg1)
```

# Correlations

```{r cc_cor,w=5,h=5, eval=TRUE,fig.cap=fig$cap("cor","Correlation on untransformed F0 data, reordered by synapse type.")}
cmatfs <- cor(fs)
corrplot(cmatfs,method="color",tl.col=ccol[ford], tl.cex=1)
```

# PCA on the Correlation Matrix

```{r pcaLog, eval=TRUE}
pcaf0 <- prcomp(featF0,scale=TRUE, center=TRUE)
pcafs <- prcomp(fs,scale=FALSE, center=FALSE)
elpcaf0 <- getElbows(pcaf0$sdev, plot=FALSE)
elpcafs <- getElbows(pcafs$sdev, plot=FALSE)
```


# Level 1: K-means++ for $K=2$. 
We run a Hierachical K-means++ for $K=2$ on the `fs` data with 4 levels.

```{r cc-kmpp,eval=TRUE,echo=TRUE}
set.seed(2^13)
L <- bhkmpp(fs,blevels=4)
```

## Within cluster correlations (Lv 1)

```{r cc-wcc1, h=8,w=8,fig.cap=fig$cap("corkp1","Within cluster correlations, clock-wise from top left, Cluster 1, Cluster 2, l2 distance between C1 and C2")}
corkp1 <- cor(fs[L[[1]] == 1,])
corkp2 <- cor(fs[L[[1]] == 2,])

par(mfrow=c(2,2))
corrplot(corkp1,method="color",tl.col=ccol[ford], tl.cex=0.8)
corrplot(corkp2,method="color",tl.col=ccol[ford], tl.cex=0.8)
corrplot(sqrt((corkp1 - corkp2)^2),method="color",tl.col=ccol[ford], tl.cex=0.8)
```

Notice that the non-synaptic markers change very little between
clusters.  Also note that the correlations between (`gad, VGAT, PV,
Gephyr`) and `VGlut1` at both times change significantly between
clusters.

## Heat maps (Lv 1):

```{r cc_agg, eval=TRUE}
## Formatting data for heatmap
aggp <- aggregate(fs,by=list(lab=L[[1]]),FUN=mean)
aggp <- as.matrix(aggp[,-1])
rownames(aggp) <- clusterFraction(L[[1]])
```

The following are heatmaps generated from clustering 
via K-means++ (at level 1)

```{r cc-kmpp-heatmapLV1,eval=FALSE,echo=FALSE,w=6,h=6,fig.cap=fig$cap("heat1", "Heatmap of the cluster means vs channels. Rows and columns are rearranged according to hclust.")}
heatmap.2(as.matrix(aggp), trace="none",col=mycol,colCol=ccol,cexRow=0.8, keysize=1,symkey=FALSE,symbreaks=FALSE,scale="none", srtCol=90,main="Heatmap of `fs` data") # 
```

```{r cc-kmpp-heatmapSortedLV1,eval=TRUE,w=6,h=6,fig.cap=fig$cap("heatfs1", "Heatmap of the cluster means vs channels. Rows and columns are rearranged according to synapse type.")}
heatmap.2(as.matrix(aggp),dendrogram='row',Colv=NA,trace="none", col=mycol,colCol=ccol[ford],cexRow=0.8, keysize=1.25,symkey=FALSE,symbreaks=FALSE,scale="none", srtCol=90,main="Heatmap of `fs` data.") 
```

Percentage of data within cluster is presented on the right side of the heatmap.


## Kernel Density Estimates of the marginals | cluster (Lv 1)

Here we look at the kernel density estimates within each cluster to
compare.

```{r synEx-kde1c2prepLV1}
df2 <- melt(as.matrix(fs))
names(df2) <- c("ind","channel","value")
df2$cluster <- L[[1]]
df2$type <- factor(rep(ffchannel,each=dim(fs)[1]),levels=levels(ffchannel))

gg2 <- ggplot(df2, aes(x=value)) + 
scale_colour_manual(values=ccol) + 
scale_x_continuous(limits=c(0,400)) +
geom_histogram(aes(y=..density..,group=channel,colour=channel),bins=250) +
geom_density(aes(group=channel, color=channel),size=1.5) +
facet_grid(channel ~ cluster, scale='free_x') + 
theme(strip.text.y=element_text(angle=0)) +
theme(strip.text.y   =element_text(angle=0),
      plot.title     =element_text(size=ts),
      axis.title.x   =element_text(size=ts),
      axis.title.y   =element_text(size=ts),
      legend.title   =element_text(size=ts),
      legend.text    =element_text(size=ts-2),
      strip.text     =element_text(size=ts), 
      legend.position='none')
```

```{r synEx-kdeLV1, eval=TRUE,echo=TRUE,cache=FALSE,w=12,h=18,fig.cap=fig$cap("gKDElv1","Kernel density estimates for each channel, on `fs` data given cluster from km++ level 1")}
print(gg2)
```

## Clusters and Spatial Location (Lv 1)
Using the location data and the results of K-means++ we show a 3d scatter
plot colored accoding to cluster.

```{r cc-kpp3dLV1}
set.seed(2^12)
s1 <- sample(dim(loc)[1],5e4)

locs1 <- loc[s1,]
locs1$cluster <- L[[1]][s1]

plot3d(locs1$V1,locs1$V2,locs1$V3,
       col=brewer.pal(4,"Set1")[order(table(L[[1]]))][locs1$cluster],
       alpha=0.75,
       xlab='x', 
       ylab='y', 
       zlab='z')

subid <- currentSubscene3d()
rglwidget(elementId="plot3dLocations", height=720, width=720)
```

# Level 2: K-means++ for $K=2$.

## Within cluster correlations (Lv 2)

```{r cc-wccLV2,h=10,w=10,fig.cap=fig$cap("corkp2","Within cluster correlations for level 2. (c11, c12, c21, c22)")}
corLV2 <- lapply(c(1:4),function(x){cor(fs[L[[2]] == x,])})

par(mfrow=c(2,2))
for(pl in corLV2){
corrplot(pl,method="color",tl.col=ccol[ford], tl.cex=0.8)
}
```

## Heat maps (Lv 2):

```{r cc-agg2, eval=TRUE}
## Formatting data for heatmap
aggp2 <- aggregate(fs,by=list(lab=L[[2]]),FUN=function(x){mean(x)}) 
aggp2 <- as.matrix(aggp2[,-1])
rownames(aggp2) <- clusterFraction(L[[2]])
```

The following are heatmaps generated from clustering 
via K-means++

```{r cc-kmpp-heatmapLV2,eval=FALSE,echo=FALSE,w=6,h=6,fig.cap=fig$cap("heat1", "Heatmap of the cluster means vs channels. Rows and columns are rearranged according to hclust.")}
heatmap.2(as.matrix(aggp2), trace="none",col=mycol,colCol=ccol,cexRow=0.8, keysize=1,symkey=FALSE,symbreaks=FALSE,scale="none", srtCol=90,main="Heatmap of `fs` data") # 
```

```{r cc-kmpp-heatmapSortedLV2,eval=TRUE,w=6,h=6,fig.cap=fig$cap("heatfs1", "Heatmap of the cluster means vs channels. Rows and columns are rearranged according to synapse type.")}
heatmap.2(as.matrix(aggp2),dendrogram='row',Colv=NA,trace="none", col=mycol,colCol=ccol[ford],cexRow=0.8, keysize=1.25,symkey=FALSE,symbreaks=FALSE,scale="none", srtCol=90,main="Heatmap of `fs` data.") 
```

Percentage of data within cluster is presented on the right side of the heatmap.


## Kernel Density Estimates of the marginals | cluster (Lv 2)

Here we look at the kernel density estimates within each cluster to
compare.

```{r synEx-kdeLV2}
df2 <- melt(as.matrix(fs))
names(df2) <- c("ind","channel","value")
df2$cluster <- L[[2]]
df2$type <- factor(rep(ffchannel,each=dim(fs)[1]),levels=levels(ffchannel))

gg3 <- ggplot(df2, aes(x=value)) + 
    scale_colour_manual(values=ccol) + 
    #scale_x_continuous(limits=c(0,400)) +
    geom_histogram(aes(y=..density..,group=channel,colour=channel),bins=250) +
    geom_density(aes(group=channel),size=.5, color='black', alpha=0.5) +
    facet_grid(channel ~ cluster, scale='free') + 
    #guides(col=guide_legend(ncol=1))
    theme(strip.text.y=element_text(angle=0),
          plot.title=element_text(size=ts),
          axis.title.x=element_text(size=ts),
          axis.title.y=element_text(size=ts),
          legend.title=element_text(size=ts),
          legend.text=element_text(size=ts-2),
          strip.text=element_text(size=ts), 
          legend.position='none')
```

```{r gg-kdeLV2, eval=TRUE,echo=TRUE,cache=FALSE,w=12,h=18,fig.cap=fig$cap("gKDE2","Kernel density estimates for each channel, on `fs` data given cluster from km++ level 2")}
print(gg3)
```

## Clusters and Spatial Location (Lv 2)
Using the location data and the results of K-means++ we show a 3d scatter
plot colored according to cluster.

```{r cc-kpp3d2}
set.seed(2^12)
s1 <- sample(dim(loc)[1],5e4)

locs2 <- loc[s1,]
locs2$cluster <- L[[2]][s1]

YlOrBr <- c("#FFFFD4", "#FED98E", "#FE9929", "#D95F0E", "#993404")
col.pal <- colorRampPalette(YlOrBr)

plot3d(locs2$V1,locs2$V2,locs2$V3,
       #col=colorpanel(8,"brown","blue")[order(table(L[[2]]))][locs2$cluster],
       col=col.pal(8)[-seq(1,8,2)][order(table(L[[2]]))][locs2$cluster],
       alpha=0.75,
       xlab='x', 
       ylab='y', 
       zlab='z'
       )

subid <- currentSubscene3d()
rglwidget(elementId="plot3dLocationsLV2", height=720, width=720)
```




# Level 3: K-means++ for $K=2$.

## Within cluster correlations (Lv 3)

```{r cc-wcc3,h=16,w=10,fig.cap=fig$cap("corkp3","Within cluster correlations for level 3. (c111, c112, c121, c122, c211, c212, c221, c222)")}
corLV3 <- lapply(c(1:8),function(x){cor(fs[L[[3]] == x,])})

par(mfrow=c(3,2))
for(pl in corLV3){
corrplot(pl,method="color",tl.col=ccol[ford], tl.cex=0.8)
}
```

## Heat maps (Lv 3):

```{r cc-agg3, eval=TRUE}
## Formatting data for heatmap
aggp3 <- aggregate(fs,by=list(lab=L[[3]]),FUN=function(x){mean(x)})
aggp3 <- as.matrix(aggp3[,-1])
rownames(aggp3) <- clusterFraction(L[[3]])
```

The following are heatmaps generated from clustering 
via K-means++

```{r cc-kmpp-heatmapLV3,eval=FALSE,echo=FALSE,w=6,h=6,fig.cap=fig$cap("heat3", "Heatmap of the cluster means vs channels. Rows and columns are rearranged according to hclust.")}
heatmap.2(as.matrix(aggp3), trace="none",col=mycol,colCol=ccol,cexRow=0.8, keysize=1,symkey=FALSE,symbreaks=FALSE,scale="none", srtCol=90,main="Heatmap of `fs` data") # 
```

```{r cc-kmpp-heatmapSortedLV3,eval=TRUE,w=6,h=6,fig.cap=fig$cap("heatfs1", "Heatmap of the cluster means vs channels. Rows and columns are rearranged according to synapse type.")}
heatmap.2(as.matrix(aggp3),dendrogram='row',Colv=NA,trace="none", col=mycol,colCol=ccol[ford],cexRow=0.8, keysize=1.25,symkey=FALSE,symbreaks=FALSE,scale="none", srtCol=90,main="Heatmap of `fs` data.") 
```

Percentage of data within cluster is presented on the right side of the heatmap.


## Kernel Density Estimates of the marginals | cluster (Lv 3)

Here we look at the kernel density estimates within each cluster to
compare.

```{r synEx-kde1c2prep3}
df3 <- melt(as.matrix(fs))
names(df3) <- c("ind","channel","value")
df3$cluster <- L[[3]]
df3$type <- factor(rep(ffchannel,each=dim(fs)[1]),levels=levels(ffchannel))

gg4 <- ggplot(df3, aes(x=value)) + 
    scale_colour_manual(values=ccol) + 
    #scale_x_continuous(limits=c(0,400)) +
    geom_histogram(aes(y=..density..,group=channel,colour=channel),bins=250) +
    geom_density(aes(group=channel),size=.5, color='black', alpha=0.5) +
    facet_grid(channel ~ cluster, scale='free') + 
    #guides(col=guide_legend(ncol=1)) + 
    theme(strip.text.y=element_text(angle=0),
          plot.title=element_text(size=ts),
          axis.title.x=element_text(size=ts),
          axis.title.y=element_text(size=ts),
          legend.title=element_text(size=ts),
          legend.text=element_text(size=ts-2),
          strip.text=element_text(size=ts), 
          legend.position='none')
```

```{r synEx-kdeLV3, eval=TRUE,echo=TRUE,cache=FALSE,w=12,h=18,fig.cap=fig$cap("gKDE3","Kernel density estimates for each channel, on `fs` data given cluster from km++ level 3")}
print(gg4)
```

## Clusters and Spatial Location (Lv 3)
Using the location data and the results of K-means++ we show a 3d scatter
plot colored according to cluster.

```{r cc-kppLV3}
set.seed(2^12)
s1 <- sample(dim(loc)[1],5e4)

locs3 <- loc[s1,]
locs3$cluster <- L[[3]][s1]

plot3d(locs3$V1,locs3$V2,locs3$V3,
       col=col.pal(16)[-seq(1,8,2)][order(table(L[[3]]))][locs3$cluster],
       alpha=0.65,
       xlab='x', 
       ylab='y', 
       zlab='z'
       )

subid <- currentSubscene3d()
rglwidget(elementId="plot3dLocationsLV3", height=720, width=720)
```







<footer>
<p> [Back to top][Introduction]</p>
</footer>
