--- 
title: "Synapse Stats: Synapse Exploration on *Log F1*"
author: "JLP"
date: '`r Sys.Date()`'
output:
  html_document:
    fig_caption: yes
    fig_height: 5
    fig_width: 5
    highlight: pygments
    keep_md: yes
    number_sections: yes
    theme: cerulean
    toc: yes
    toc_depth: 3
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
---
```{r render, eval=FALSE, echo=FALSE}
require(rmarkdown)
rm(list=ls()); 
rmarkdown::render(grep("^SynapseExplorationLogF1.Rmd", dir(), value=TRUE))
system('open SynapseExplorationLogF1.html')
```

```{r setup,include=FALSE,results='asis',message=FALSE,warning=FALSE}
### Library calls here.
require(Matrix)
require(colorRamps)
require(RColorBrewer)
require(hexbin)
require(corrplot)
require(gplots)
require(dplyr)
require(reshape2)
require(ggplot2)
require(gridExtra)
require(lattice)
require(parallel)
require(data.table)
require(fpc)
require(repr)
require(mclust)
require(doMC)
require(rgl)
require(rglwidget)

source("./bic.r")
source("./clusterFraction.r")
source("./kmpp.r")
#source("./getElbows.r")
source("http://www.cis.jhu.edu/~parky/Synapse/getElbows.R")
registerDoMC(6)

library(rmarkdown)
library(knitr)
```
```{r knitr-setup, include=FALSE, results='asis'}
### The following options and figure numbering functions
### were setup by Youngser Park
knitr::opts_chunk$set(cache=TRUE, autodep=TRUE)
dep_auto() # figure out dependencies automatically
opts_chunk$set(cache=FALSE, echo=TRUE, 
               warning=FALSE, message=FALSE, 
               comment="#", fig.path='../Figures/SynapseExplorationLogF1_figure/',
               dpi=227,dev=c('png','pdf'))

opts_knit$set(aliases = c(h = 'fig.height', w = 'fig.width', 
                          cap='fig.cap', scap='figscap'))

opts_knit$set(eval.after = c('fig.cap','fig.scap'))

knit_hooks$set(document = function(x) {
  gsub('(\\\\end\\{knitrout\\}[\n]+)', '\\1\\\\noindent ', x)
  }
)
#opts_knit$set(animation.fun = hook_scianimator)

knit_hooks$set(plot = function(x, options) {
  paste('<figure><img src="',
        opts_knit$get('base.url'), 
        paste(x, collapse = '.'),
             '"><figcaption>', options$fig.cap, '</figcaption></figure>',
             sep = '')
 })
 
 fn = local({
   i = 0
   function(x) {
     i <<- i + 1
     paste('Figure ', i, ': ', x, sep = '')
   }
 })

 fig <- local({
    i <- 0
    ref <- list()
    list(
        cap=function(refName, text) {
            i <<- i + 1
            ref[[refName]] <<- i
            paste("<b>Figure ", i, ": ", text, "</b><br><br>", sep="")
        },
        ref=function(refName) {
            ref[[refName]]
        })
})
```

[Homepage](http://docs.neurodata.io/synaptome-stats/)  
The formatted source code for this file is 
[here](https://github.com/neurodata/synaptome-stats/blob/gh-pages/Code/SynapseExploration.Rmd).  
And a [raw version here](https://raw.githubusercontent.com/neurodata/synaptome-stats/gh-pages/Code/SynapseExploration.Rmd).    
Previous work by Youngser Park can be found [here](http://www.cis.jhu.edu/~parky/Synapse/synapse.html).  


```{r cc-data, eval=TRUE, echo=FALSE}
featFull <- fread("../data/synapsinR_7thA.tif.Pivots.txt.2011Features.txt",
                  showProgress=FALSE)
locFull <- fread("../data/synapsinR_7thA.tif.Pivots.txt",showProgress=FALSE)

### Setting a seed and creating an index vector
### to select half of the data
set.seed(2^10)
half1 <- sample(dim(featFull)[1],dim(featFull)[1]/2)
half2 <- setdiff(1:dim(featFull)[1],half1)

feat <- featFull[half1,]
loc <- locFull[half1,]

## Setting the channel names
channel <- c('Synap_1','Synap_2','VGlut1_t1','VGlut1_t2','VGlut2','Vglut3',
              'psd','glur2','nmdar1','nr2b','gad','VGAT',
              'PV','Gephyr','GABAR1','GABABR','CR1','5HT1A',
              'NOS','TH','VACht','Synapo','tubuli','DAPI')

## Setting the channel types
channel.type <- c('ex.pre','ex.pre','ex.pre','ex.pre','ex.pre','in.pre.small',
                  'ex.post','ex.post','ex.post','ex.post','in.pre','in.pre',
                  'in.pre','in.post','in.post','in.post','in.pre.small','other',
                  'ex.post','other','other','ex.post','none','none')

nchannel <- length(channel)
nfeat <- ncol(feat) / nchannel

## Createing factor variables for channel and channel type sorted properly
ffchannel <- (factor(channel.type,
    levels= c("ex.pre","ex.post","in.pre","in.post","in.pre.small","other","none")
    ))
fchannel <- as.numeric(factor(channel.type,
    levels= c("ex.pre","ex.post","in.pre","in.post","in.pre.small","other","none")
    ))
ford <- order(fchannel)


## Setting up colors for channel types
Syncol <- c("#197300","#5ed155","#660000","#cc0000","#ff9933","mediumblue","gold")
ccol <- Syncol[fchannel]

exType <- factor(c(rep("ex",11),rep("in",6),rep("other",7)),ordered=TRUE)
exCol<-exType;levels(exCol) <- c("#197300","#990000","mediumblue");
exCol <- as.character(exCol)

fname <- as.vector(sapply(channel,function(x) paste0(x,paste0("F",0:5))))
names(feat) <- fname
fcol <- rep(ccol, each=6)
mycol <- colorpanel(100, "purple", "black", "green")
mycol2 <- matlab.like(nchannel)
```

```{r cc-datatrans, eval=TRUE, echo=FALSE}
f <- lapply(1:6,function(x){seq(x,ncol(feat),by=nfeat)})
featF <- lapply(f,function(x){subset(feat,select=x)})

DAT <- featF[[2]]
f11e3 <- 1e3*data.table(apply(X=DAT, 2, function(x){((x-min(x))/(max(x)-min(x)))}))

fs <- f11e3

### Taking log_10 on data with 0's removed
ans <- apply(DAT, 1, function(row){ any(row == 0)})

logF1 <- log10(DAT[!ans,])
slogF1 <- logF1[,lapply(.SD,scale, center=TRUE,scale=TRUE)]
```

We now have the following data sets:

- `featF1`: The feature vector looking only at the integrated brightness features.
- `fs`:  The feature vector scaled between $[0,1000]$.
- `logF1`: The feature vector, with 0's removed, then $log_{10}$ is applied. 
- `slogF1`: The feature vector, with 0's removed, then $log_{10}$, then scaled by
  subtracting the mean and dividing by the sample standard deviation.


# Level 0 

```{r cc-dat0,eval=TRUE,echo=TRUE}
dat <- logF1
```

## Heat maps (Lv 0):

```{r cc_agg0, eval=TRUE}
## Formatting data for heatmap
aggp <- apply(dat, 2, mean)
aggp <- t(cbind(aggp, aggp))[, ford]
```

The following are heatmaps generated from clustering 
via K-means++ (at level 1)


```{r cc-kmpp-heatmapSortedLV0,eval=TRUE,w=6,h=6,fig.cap=fig$cap("heatlogF11", "Heatmap of the cluster means vs channels. Rows and columns are rearranged according to synapse type.")}
heatmap.2(as.matrix(aggp),dendrogram='none',Colv=NA,trace="none", 
          col=mycol,colCol=ccol[ford],cexRow=0.8, keysize=1.25,symkey=FALSE,
          symbreaks=FALSE,scale="none", srtCol=90,main="Heatmap of `logF1` data.", 
          labRow = "") 
```

Percentage of data within cluster is presented on the right side of the heatmap.


## Jittered scatter plot: Lv 0

```{r cc-kmpp0,eval=TRUE,echo=FALSE}
set.seed(2^13)
L <- bhkmpp(dat,blevels=4)
```

```{r ggJitter}
set.seed(1024)
s2 <- sample(dim(dat)[1], 1e4)
ggJdat <- data.table(cbind(stack(dat[s2]),L[s2]))
ggJdat$ind <- as.factor(ggJdat$ind)
levels(ggJdat$ind) <- names(dat)[ford]

ggJ0 <- 
  ggplot(data = ggJdat, aes(x = ind, y = values)) +
  geom_point(alpha=0.75) + 
  geom_jitter(width = 1) + 
  geom_boxplot(alpha =0.35, outlier.color = 'NA') + 
  theme(axis.title.x = element_blank()) + 
  theme(axis.text.x = element_text(color = ccol[ford], 
                                   angle=45,
                                   vjust = 0.5))
```

```{r jitter0, h=6, w=14, fig.cap=fig$cap("jitter1","Scatter Plot Level 0 ")}
print(ggJ0)
```

The above scatter plot is a random sample of the data points. 

## Correlations: Lv 0

```{r cc_cor,w=5,h=5, eval=TRUE,fig.cap=fig$cap("cor","Correlation on untransformed F1 data, reordered by synapse type.")}
cmat <- cor(dat)[ford, ford]
corrplot(cmat,method="color",tl.col=ccol[ford], tl.cex=1)
```


# Level 1: K-means++ for $K=2$. 
We run a Hierachical K-means++ for $K=2$ on the `logF1` data with 4 levels.

```{r cc-kmpp,eval=TRUE,echo=TRUE}
set.seed(2^13)
L <- bhkmpp(dat,blevels=4)
```

## Heat maps (Lv 1):

```{r cc_agg, eval=TRUE}
## Formatting data for heatmap
aggp <- aggregate(dat,by=list(lab=L[[1]]),FUN=mean)
aggp <- as.matrix(aggp[,-1])[, ford]
rownames(aggp) <- clusterFraction(L[[1]])
```

The following are heatmaps generated from clustering 
via K-means++ (at level 1)

```{r cc-kmpp-heatmapLV1,eval=FALSE,echo=FALSE,w=6,h=6,fig.cap=fig$cap("heat1", "Heatmap of the cluster means vs channels. Rows and columns are rearranged according to hclust.")}
heatmap.2(as.matrix(aggp), trace="none",col=mycol,colCol=ccol,cexRow=0.8, keysize=1,symkey=FALSE,symbreaks=FALSE,scale="none", srtCol=90,main="Heatmap of `logF1` data") # 
```

```{r cc-kmpp-heatmapSortedLV1,eval=TRUE,w=6,h=6,fig.cap=fig$cap("heatlogF11", "Heatmap of the cluster means vs channels. Rows and columns are rearranged according to synapse type.")}
heatmap.2(as.matrix(aggp),dendrogram='row',Colv=NA,trace="none", col=mycol,colCol=ccol[ford],cexRow=0.8, keysize=1.25,symkey=FALSE,symbreaks=FALSE,scale="none", srtCol=90,main="Heatmap of `logF1` data.") 
```

Percentage of data within cluster is presented on the right side of the heatmap.


## Jittered scatter plot: Lv 1


```{r ggJitter1}
ggCol <- brewer.pal(4,"Set1")[order(table(L[[1]]))]

cf1 <- data.frame(cf = clusterFraction(L[[1]]))

ggJ1 <- 
  ggplot(data = ggJdat, aes(x = ind, y = values, 
                         color = as.factor(lv1))) +
  scale_color_manual(values=ggCol, name="Cluster") + 
  geom_point(alpha=0.25, position=position_jitterdodge()) + 
  geom_boxplot(alpha =0.35, outlier.color = 'NA') + 
  annotate("text", x = levels(ggJdat$ind)[c(2,20)], y = 1.15*max(ggJdat$values), 
           label= cf1[1:2,]) + 
  theme(axis.title.x = element_blank()) + 
  theme(axis.text.x = element_text(color = ccol[ford], 
                                   angle=45,
                                   vjust = 0.5))
```

```{r jitter1, h=6, w=14, fig.cap=fig$cap("jitter1","Scatter Plot Level 1 ")}
print(ggJ1)
```

## Within cluster correlations (Lv 1)

```{r cc-wcc1, h=8,w=8,fig.cap=fig$cap("corkp1","Within cluster correlations, clock-wise from top left, Cluster 1, Cluster 2, difference C1 - C2")}
corkp1 <- cor(dat[L[[1]] == 1,])
corkp2 <- cor(dat[L[[1]] == 2,])
difCor12 <- (corkp1 - corkp2)

layout(matrix(c(1,2,3,3), 2, 2, byrow=TRUE))
corrplot(corkp1,method="color",tl.col=ccol[ford], tl.cex=0.8, mar=c(0,0,3,0))
title("Cluster 1")
corrplot(corkp2,method="color",tl.col=ccol[ford], tl.cex=0.8, mar=c(0,0,3,0))
title("Cluster 2")
corrplot(difCor12,is.corr=FALSE,method="color",
         tl.col=ccol[ford], tl.cex=0.8, 
         mar=c(0,0,3,0),
         col=colorRampPalette(c("#998ec3","white","darkorange"))(50))
title("Difference(1,2)")
```


## Clusters and Spatial Location (Lv 1)
Using the location data and the results of K-means++ we show a 3d scatter
plot colored accoding to cluster.

```{r cc-kpp3dLV1}
set.seed(2^12)
s1 <- sample(dim(loc)[1],5e4)

locs1 <- loc[s1,]
locs1$cluster <- L[[1]][s1]

plot3d(locs1$V1,locs1$V2,locs1$V3,
       col=brewer.pal(4,"Set1")[order(table(L[[1]]))][locs1$cluster],
       alpha=0.75,
       xlab='x', 
       ylab='y', 
       zlab='z')

subid <- currentSubscene3d()
rglwidget(elementId="plot3dLocations", height=720, width=720)
```

# Level 2: K-means++ for $K=2$.


## Heat maps (Lv 2):

```{r cc-agg2, eval=TRUE}
## Formatting data for heatmap
aggp2 <- aggregate(dat,by=list(lab=L[[2]]),FUN=function(x){mean(x)}) 
aggp2 <- as.matrix(aggp2[,-1])[, ford]
rownames(aggp2) <- clusterFraction(L[[2]])
```

The following are heatmaps generated from clustering 
via K-means++

```{r cc-kmpp-heatmapLV2,eval=FALSE,echo=FALSE,w=6,h=6,fig.cap=fig$cap("heat1", "Heatmap of the cluster means vs channels. Rows and columns are rearranged according to hclust.")}
heatmap.2(as.matrix(aggp2), trace="none",col=mycol,colCol=ccol,cexRow=0.8, keysize=1,symkey=FALSE,symbreaks=FALSE,scale="none", srtCol=90,main="Heatmap of `logF1` data") # 
```

```{r cc-kmpp-heatmapSortedLV2,eval=TRUE,w=6,h=6,fig.cap=fig$cap("heat1", "Heatmap of the cluster means vs channels. Rows and columns are rearranged according to synapse type.")}
heatmap.2(as.matrix(aggp2),dendrogram='row',Colv=NA,trace="none", col=mycol,colCol=ccol[ford],cexRow=0.8, keysize=1.25,symkey=FALSE,symbreaks=FALSE,scale="none", srtCol=90,main="Heatmap of `logF1` data.") 
```

Percentage of data within cluster is presented on the right side of the heatmap.


## Jittered scatter plot: Lv 2


```{r ggJitter2}
ggCol <- brewer.pal(8,"Set1")[order(table(L[[2]]))]

cf2 <- data.frame(cf = clusterFraction(L[[2]]))

ggJ2 <- 
  ggplot(data = ggJdat, aes(x = ind, y = values, 
                         color = as.factor(lv2))) +
  scale_color_manual(values=ggCol, name = 'Cluster') + 
  geom_point(alpha=0.25, position=position_jitterdodge()) + 
  geom_boxplot(alpha =0.35, outlier.color = 'NA') + 
  annotate("text", x = levels(ggJdat$ind)[c(2,8,14,20)], y = 1.15*max(ggJdat$values), 
           label= cf2[1:4,]) + 
  theme(axis.title.x = element_blank()) + 
  theme(axis.text.x = element_text(color = ccol[ford], 
                                   angle=45,
                                   vjust = 0.5))
```

```{r jitter2, h=6, w=14, fig.cap=fig$cap("jitter2","Scatter Plot Level 2 ")}
print(ggJ2)
```

The fraction of data points within each cluster are given at the top of
the plot window. 


## Within cluster correlations (Lv 2)

```{r cc-wccLV2,h=12,w=7,fig.cap=fig$cap("corkp2","Within cluster correlations for level 2. (c11, c12, c21, c22) with differences")}
corLV2 <- lapply(c(1:4),function(x){cor(dat[L[[2]] == x,])})

difCor1112 <- ((corLV2[[1]] - corLV2[[2]]))
difCor2122 <- ((corLV2[[3]] - corLV2[[4]]))

layout(matrix(c(1,2,3,3,4,5,6,6), 4, 2, byrow=TRUE))
corrplot(corLV2[[1]],method="color",tl.col=ccol[ford], tl.cex=0.8, 
         mar=c(0,0,3,0))
title("Cluster 1")
corrplot(corLV2[[2]],method="color",tl.col=ccol[ford], tl.cex=0.8, 
         mar=c(0,0,3,0))
title("Cluster 2")
corrplot(difCor1112, method="color", tl.col=ccol[ford], 
         tl.cex=0.8,
         mar = c(0,0,3,0),
         cl.lim = c(min(difCor1112,difCor2122),max(difCor1112,difCor2122)),
         col=colorRampPalette(c("#998ec3", 
                                "white",
                                "darkorange"))(100))
title("Difference(1,2)")
corrplot(corLV2[[3]],method="color",tl.col=ccol[ford], tl.cex=0.8, 
         mar=c(0,0,3,0))
title("Cluster 3")
corrplot(corLV2[[4]],method="color",tl.col=ccol[ford], tl.cex=0.8, 
         mar=c(0,0,3,0))
title("Cluster 4")
corrplot(difCor2122, method="color", tl.col=ccol[ford], 
         tl.cex=0.8,
         mar=c(0,0,3,0),
         cl.lim = c(min(difCor1112,difCor2122),max(difCor1112,difCor2122)),
         col=colorRampPalette(c("#998ec3", 
                                "white",
                                "darkorange"))(100))
title("Difference(3,4)")
```

## Clusters and Spatial Location (Lv 2)
Using the location data and the results of K-means++ we show a 3d scatter
plot colored according to cluster.

```{r cc-kpp3d2}
set.seed(2^12)
s1 <- sample(dim(loc)[1],5e4)

locs2 <- loc[s1,]
locs2$cluster <- L[[2]][s1]

YlOrBr <- c("#FFFFD4", "#FED98E", "#FE9929", "#D95F0E", "#993404")
col.pal <- colorRampPalette(YlOrBr)

plot3d(locs2$V1,locs2$V2,locs2$V3,
       #col=colorpanel(8,"brown","blue")[order(table(L[[2]]))][locs2$cluster],
       col=col.pal(8)[-seq(1,8,2)][order(table(L[[2]]))][locs2$cluster],
       alpha=0.75,
       xlab='x', 
       ylab='y', 
       zlab='z'
       )

subid <- currentSubscene3d()
rglwidget(elementId="plot3dLocationsLV2", height=720, width=720)
```




# Level 3: K-means++ for $K=2$.


## Heat maps (Lv 3):

```{r cc-agg3, eval=TRUE}
## Formatting data for heatmap
aggp3 <- aggregate(dat,by=list(lab=L[[3]]),FUN=function(x){mean(x)})
aggp3 <- as.matrix(aggp3[,-1])[, ford]
rownames(aggp3) <- clusterFraction(L[[3]])
```

The following are heatmaps generated from clustering 
via K-means++

```{r cc-kmpp-heatmapLV3,eval=FALSE,echo=FALSE,w=6,h=6,fig.cap=fig$cap("heat3", "Heatmap of the cluster means vs channels. Rows and columns are rearranged according to hclust.")}
heatmap.2(as.matrix(aggp3), trace="none",col=mycol,colCol=ccol,cexRow=0.8, keysize=1,symkey=FALSE,symbreaks=FALSE,scale="none", srtCol=90,main="Heatmap of `logF1` data") # 
```

```{r cc-kmpp-heatmapSortedLV3,eval=TRUE,w=6,h=6,fig.cap=fig$cap("heatlogF11", "Heatmap of the cluster means vs channels. Rows and columns are rearranged according to synapse type.")}
heatmap.2(as.matrix(aggp3),dendrogram='row',Colv=NA,trace="none", col=mycol,colCol=ccol[ford],cexRow=0.8, keysize=1.25,symkey=FALSE,symbreaks=FALSE,scale="none", srtCol=90,main="Heatmap of `logF1` data.") 
```

Percentage of data within cluster is presented on the right side of the heatmap.


## Jittered scatter plot: Lv 3


```{r ggJitter3}
ggCol <- brewer.pal(8,"Set1")[order(table(L[[3]]))]
cf3 <- data.frame(cf = clusterFraction(L[[3]]))

ggJ3 <- 
  ggplot(data = ggJdat, aes(x = ind, y = values, 
                         color = as.factor(lv3))) +
  scale_color_manual(values=ggCol, name = "Cluster") + 
  geom_point(alpha=0.25, position=position_jitterdodge()) + 
  geom_boxplot(alpha =0.35, outlier.color = 'NA') + 
  annotate("text", x = levels(ggJdat$ind)[seq(2,22, length=8)], y = 1.05*max(ggJdat$values), 
           label= cf3[1:8,]) + 
  #geom_jitter(width=2) + 
  theme(axis.title.x = element_blank()) + 
  theme(axis.text.x = element_text(color = ccol[ford], 
                                   angle=45,
                                   vjust = 0.5))
```

```{r jitter3, h=6, w=17, fig.cap=fig$cap("jitter3","Scatter Plot Level 3 ")}
print(ggJ3)
```

## Within cluster correlations (Lv 3)

```{r cc-wcc3,h=17,w=5,fig.cap=fig$cap("corkp3","Within cluster correlations for level 3. (c111, c112, c121, c122, c211, c212, c221, c222)")}
corLV3 <- lapply(c(1:8),function(x){cor(dat[L[[3]] == x,])[ford, ford]})

difCor1 <- (corLV3[[1]] - corLV3[[2]])
difCor2 <- (corLV3[[3]] - corLV3[[4]])
difCor3 <- (corLV3[[5]] - corLV3[[6]])
difCor4 <- (corLV3[[7]] - corLV3[[8]])
M <- max(difCor1, difCor2, difCor3, difCor4)
m <- min(difCor1, difCor2, difCor3, difCor4)

layout(matrix(c(1, 2, 3, 3,
                4, 5, 6, 6, 
                7, 8, 9, 9,
                10, 11, 12, 12), 8,2, byrow=TRUE))

corrplot(corLV3[[1]],method="color",tl.col=ccol[ford], tl.cex=0.8,
         mar=c(0,0,3,0))
title('Cluster 1')
corrplot(corLV3[[2]],method="color",tl.col=ccol[ford], tl.cex=0.8,
         mar=c(0,0,3,0))
title('Cluster 2')
corrplot(difCor1,method="color",tl.col=ccol[ford], tl.cex=0.8,
         cl.lim=c(m,M), 
         mar=c(0,0,3,0),
         col=colorRampPalette(c("#998ec3",
                                "white",
                                "darkorange"))(50))
title('Difference(1,2)')
corrplot(corLV3[[3]],method="color",tl.col=ccol[ford], tl.cex=0.8, 
         mar=c(0,0,3,0))
title('Cluster 3')
corrplot(corLV3[[4]],method="color",tl.col=ccol[ford], tl.cex=0.8, 
         mar=c(0,0,3,0))
title('Cluster 4')
corrplot(difCor2,method="color",tl.col=ccol[ford], tl.cex=0.8, 
         cl.lim= c(m,M),
         mar=c(0,0,3,0),
         col=colorRampPalette(c("#998ec3",
                                "white",
                                "darkorange"))(50))
title('Difference(3,4)')
corrplot(corLV3[[5]],method="color",tl.col=ccol[ford], tl.cex=0.8, 
         mar=c(0,0,3,0))
title('Cluster 5')
corrplot(corLV3[[6]],method="color",tl.col=ccol[ford], tl.cex=0.8, 
         mar=c(0,0,3,0))
title('Cluster 6')
corrplot(difCor3,method="color",tl.col=ccol[ford], tl.cex=0.8,
         cl.lim= c(m,M),
         mar=c(0,0,3,0),
         col=colorRampPalette(c("#998ec3",
                                "white",
                                "darkorange"))(50))
title('Difference(5,6)')
corrplot(corLV3[[7]],method="color",tl.col=ccol[ford], tl.cex=0.8, 
         mar=c(0,0,3,0))
title('Cluster 7')
corrplot(corLV3[[8]],method="color",tl.col=ccol[ford], tl.cex=0.8, 
         mar=c(0,0,3,0))
title('Cluster 8')
corrplot(difCor4,method="color",tl.col=ccol[ford], tl.cex=0.8,
         cl.lim= c(m,M),
         mar=c(0,0,3,0),
         col=colorRampPalette(c("#998ec3",
                                "white",
                                "darkorange"))(50))
title('Difference(7,8)')
```
## Clusters and Spatial Location (Lv 3)
Using the location data and the results of K-means++ we show a 3d scatter
plot colored according to cluster.

```{r cc-kppLV3}
set.seed(2^12)
s1 <- sample(dim(loc)[1],5e4)

locs3 <- loc[s1,]
locs3$cluster <- L[[3]][s1]

plot3d(locs3$V1,locs3$V2,locs3$V3,
       col=col.pal(16)[-seq(1,8,2)][order(table(L[[3]]))][locs3$cluster],
       alpha=0.65,
       xlab='x', 
       ylab='y', 
       zlab='z'
       )

subid <- currentSubscene3d()
rglwidget(elementId="plot3dLocationsLV3", height=720, width=720)
```


<footer>
<p> [Back to top][Level 0]</p>
</footer>
