---
title: "Synapse Stats: Synapse Exploration on *Lin F0*"
author: "JLP"
date: '`r Sys.Date()`'
output:
  html_document:
    fig_caption: yes
    fig_height: 5
    fig_width: 5
    highlight: pygments
    keep_md: yes
    number_sections: yes
    theme: cerulean
    toc: yes
    toc_depth: 3
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
---
```{r render, eval=FALSE, echo=FALSE}
require(rmarkdown)
rm(list=ls()); 
rmarkdown::render(grep("^SynapseExploration.Rmd", dir(), value=TRUE))
system('open SynapseExploration.html')
system('say -r 200 Your R-script has completed')
```

```{r setup,include=FALSE,results='asis',message=FALSE,warning=FALSE}
### Library calls here.
require(Matrix)
require(MASS)
require(colorRamps)
require(RColorBrewer)
require(hexbin)
require(corrplot)
require(gplots)
require(dplyr)
require(reshape2)
require(ggplot2)
require(gridExtra)
require(lattice)
require(parallel)
require(data.table)
require(fpc)
require(repr)
require(mclust)
require(doMC)
require(rgl)
require(rglwidget)
require(deldir)
source("./bic.r")
source("./clusterFraction.r")
source("./kmpp.r")
#source("./getElbows.r")
source("http://www.cis.jhu.edu/~parky/Synapse/getElbows.R")
registerDoMC(6)

library(rmarkdown)
library(knitr)
```
```{r knitr-setup, include=FALSE, results='asis'}
### The following options and figure numbering functions
### were setup by Youngser Park
knitr::opts_chunk$set(cache=TRUE, autodep=TRUE)
dep_auto() # figure out dependencies automatically
opts_chunk$set(cache=FALSE, echo=TRUE, 
               warning=FALSE, message=FALSE, fig.show='hold',
               comment="#", fig.path='../Figures/SynapseExploration_figure/',
               dpi=227,dev=c('png','pdf'))

opts_knit$set(aliases = c(h = 'fig.height', w = 'fig.width', 
                          cap='fig.cap', scap='figscap'))

opts_knit$set(eval.after = c('fig.cap','fig.scap'))

knit_hooks$set(document = function(x) {
  gsub('(\\\\end\\{knitrout\\}[\n]+)', '\\1\\\\noindent ', x)
  }
)
#opts_knit$set(animation.fun = hook_scianimator)

knit_hooks$set(plot = function(x, options) {
  paste('<figure><img src="',
        opts_knit$get('base.url'), 
        paste(x, collapse = '.'),
             '"><figcaption>', options$fig.cap, '</figcaption></figure>',
             sep = '')
 })
 
 fn = local({
   i = 0
   function(x) {
     i <<- i + 1
     paste('Figure ', i, ': ', x, sep = '')
   }
 })

 fig <- local({
    i <- 0
    ref <- list()
    list(
        cap=function(refName, text) {
            i <<- i + 1
            ref[[refName]] <<- i
            paste("<b>Figure ", i, ": ", text, "</b><br><br>", sep="")
        },
        ref=function(refName) {
            ref[[refName]]
        })
})
```

[Homepage](http://docs.neurodata.io/synaptome-stats/)  
The formatted source code for this file is 
[here](https://github.com/neurodata/synaptome-stats/blob/gh-pages/Code/SynapseExploration.Rmd).  
And a [raw version here](https://raw.githubusercontent.com/neurodata/synaptome-stats/gh-pages/Code/SynapseExploration.Rmd).    
Previous work by Youngser Park can be found [here](http://www.cis.jhu.edu/~parky/Synapse/synapse.html).  

```{r cc-data, eval=TRUE, echo=FALSE}
featFull <- fread("../data/synapsinR_7thA.tif.Pivots.txt.2011Features.txt",
                  showProgress=FALSE)
locFull <- fread("../data/synapsinR_7thA.tif.Pivots.txt",showProgress=FALSE)

### Setting a seed and creating an index vector
### to select half of the data
set.seed(2^10)
half1 <- sample(dim(featFull)[1],dim(featFull)[1]/2)
half2 <- setdiff(1:dim(featFull)[1],half1)

feat <- featFull[half1,]
feat2 <- featFull[half2,]
loc <- locFull[half1,]

## Setting the channel names
channel <- c('Synap_1','Synap_2','VGlut1_t1','VGlut1_t2','VGlut2','Vglut3',
              'psd','glur2','nmdar1','nr2b','gad','VGAT',
              'PV','Gephyr','GABAR1','GABABR','CR1','5HT1A',
              'NOS','TH','VACht','Synapo','tubuli','DAPI')

## Setting the channel types
channel.type <- c('ex.pre','ex.pre','ex.pre','ex.pre','ex.pre','in.pre.small',
                  'ex.post','ex.post','ex.post','ex.post','in.pre','in.pre',
                  'in.pre','in.post','in.post','in.post','in.pre.small','other',
                  'ex.post','other','other','ex.post','none','none')

nchannel <- length(channel)
nfeat <- ncol(feat) / nchannel

## Createing factor variables for channel and channel type sorted properly
ffchannel <- (factor(channel.type,
    levels= c("ex.pre","ex.post","in.pre","in.post","in.pre.small","other","none")
    ))
fchannel <- as.numeric(factor(channel.type,
    levels= c("ex.pre","ex.post","in.pre","in.post","in.pre.small","other","none")
    ))
ford <- order(fchannel)


## Setting up colors for channel types
Syncol <-  c("#197300","#5ed155","#660000","#cc0000","#ff9933","#0000cd","#ffd700")
Syncol3 <- c("#197300","#197300","#cc0000","#cc0000","#0000cd","#0000cd","#0000cd")
ccol <- Syncol[fchannel]
ccol3 <- Syncol3[fchannel]

exType <- factor(c(rep("ex",11),rep("in",6),rep("other",7)),ordered=TRUE)
exCol<-exType;levels(exCol) <- c("#197300","#990000","mediumblue");
exCol <- as.character(exCol)

fname <- as.vector(sapply(channel,function(x) paste0(x,paste0("F",0:5))))
names(feat) <- names(feat2) <- fname
fcol <- rep(ccol, each=6)
#mycol <- colorpanel(100, "purple", "black", "green")
mycol <- colorpanel(100, "black", "pink")
mycol2 <- matlab.like(nchannel)
```

```{r cc-datatrans, eval=TRUE, echo=FALSE}
f <- lapply(1:6,function(x){seq(x,ncol(feat),by=nfeat)})
f2 <- lapply(1:6,function(x){seq(x,ncol(feat2),by=nfeat)})
featF <- lapply(f,function(x){subset(feat,select=x)})
featF2 <- lapply(f2,function(x){subset(feat2,select=x)})

featF0 <- featF[[1]]
featF02 <- featF2[[1]]
f01e3 <- 1e3*data.table(apply(X=featF0, 2, function(x){((x-min(x))/(max(x)-min(x)))}))
f01e32 <- 1e3*data.table(apply(X=featF02, 2, function(x){((x-min(x))/(max(x)-min(x)))}))

fs <- f01e3
fs2 <- f01e32
```

# Definition of Markers

From email correspondence with Kristina Micheva we have the following
definitions of the given markers and their corresponding class colors used
throughout this document. The abbreviations are presented as they were
given. 

> On Feb 8, 2016, at 2:00 PM, Kristina Micheva <kmicheva@stanford.edu> wrote:

* <FONT COLOR=#197300>_Excitatory presynaptic: 'Synap', 'Synap', 'VGlut1', 'VGlut1', 'VGlut2'_</FONT>,
* <FONT COLOR=#5ed155>_Excitatory postsynaptic: 'psd', 'glur2', 'nmdar1', 'nr2b', 'NOS', 'Synapo'_</FONT> (but further away than PSD, gluR2, nmdar1 and nr2b)
* <FONT COLOR=#660000>_Inhibitory presynaptic: 'gad', 'VGAT', 'PV'_</FONT>,
* <FONT COLOR=#ff3333>_Inhibitory postsynaptic: 'Gephyr', 'GABAR1', 'GABABR', 'NOS'_</FONT>,
* <FONT COLOR=#ff9933>_At a very small number of inhibitory: 'Vglut3' (presynaptic), 'CR1'(presynaptic)_</FONT>,
* <FONT COLOR="mediumblue">_Other synapses:'5HT1A', 'TH', 'VACht'_</FONT>,
* <FONT COLOR="gold">_Not at synapses: 'tubuli', 'DAPI'_</FONT>.

Notes from correspondence :
- `Synap` is listed twice, this is not a mistake as
it was used twice with different antibodies -- I was not told which.
- `VGlut1` is also listed twice.  It was measured at two different times
  using the same antibody.
- `NOS` is listed in two different categories and the analysis so far
  does not allow that. It will be considered as **Excitatory
  postsynaptic** from here on.


We will be using the data scaled to $[0,1000]$ for this exploration. 

- `fs`:  The feature vector scaled between $[0,1000]$.

# Level 0 

Level 0 refers to the un-clustered data, that is all of the data lives
in one cluster.

```{r cc-dat0,eval=TRUE,echo=TRUE}
dat <- fs
dat2 <- fs2
```

## Heat maps (Lv 0):

```{r cc_agg0, eval=TRUE}
## Formatting data for heatmap by taking the column means.
## heatmap.2 requires at least two rows and columns. 
aggp <- apply(dat, 2, mean)
aggp <- t(cbind(aggp, aggp))[, ford]
```

The following is a heatmap showing the averages of the marker values
over the entire dataset.  The reader should notice that the markers
believed to be excitatory have higher means than do markersthe difference
between groups of markers as compared with their believed classes.

```{r cc-kmpp-heatmapSortedLV0,eval=TRUE,w=6,h=5,fig.cap=fig$cap("heatfs1", "Heatmap of the marker means. Columns are rearranged according to synapse type.")}
heatmap.2(as.matrix(aggp),dendrogram='none',Colv=NA,trace="none", 
          col=mycol,colCol=ccol[ford],cexRow=0.8, keysize=1.25,symkey=FALSE,
          symbreaks=FALSE,scale="none", srtCol=90,main="Heatmap of `fs` data.", 
          labRow = "") 
```


## Jittered scatter plot: Lv 0

```{r cc-kmpp0,eval=TRUE,echo=FALSE}
set.seed(2^13)
L <- bhkmpp(dat,blevels=4)
```

```{r ggJitter}
set.seed(1024)
s2 <- sample(dim(dat)[1], 1e4)
ggJdat <- data.table(cbind(stack(dat[s2]),L[s2]))
ggJdat$ind <- factor(ggJdat$ind, ordered=TRUE, levels=names(dat))

ggJ0 <- 
  ggplot(data = ggJdat, aes(x = ind, y = values)) +
  geom_point(alpha=0.25) + 
  geom_jitter(width = 1) + 
  geom_boxplot(alpha =0.35, outlier.color = 'NA') + 
  theme(axis.title.x = element_blank()) + 
  theme(axis.text.x = element_text(color = ccol[ford], 
                                   angle=45,
                                   vjust = 0.5))
```

```{r jitter0, h=6, w=14, fig.cap=fig$cap("jitter1","Scatter Plot Level 0 ")}
print(ggJ0)
```

The above scatter plot is a random sample of the data points. 

## Correlations: Lv 0

```{r cc_cor,w=5,h=5, eval=TRUE,fig.cap=fig$cap("cor","Correlation on untransformed F0 data, reordered by synapse type.")}
cmatfs <- cor(fs)[ford, ford]
corrplot(cmatfs,method="color",tl.col=ccol[ford], tl.cex=1)
```

### PCA of the within cluster correlation matrices at level 0.
```{r cc-pca0}
pcaL0 <- prcomp(cmatfs, center=TRUE, scale=TRUE)
```

#### PCA Lv0
```{r cc_3dCorr0,echo=TRUE, eval=TRUE, fig.cap=fig$cap("rgl0","PCA 1-3 on untransformed correlation matrix")}
pca0 <- data.frame(pcaL0$x)
rgl::plot3d(pca0[,1],pca0[,2],pca0[,3],type='s',col=ccol3[ford], size=1,
            xlab = "PC1", ylab = "PC2", zlab = "PC3")
abclines3d(0, 0, 0, a = diag(3), col = "gray")
rgl::rgl.texts(pca0[,1],pca0[,2],pca0[,3],abbreviate(rownames(pca0)), col=ccol3[ford], adj=c(0,2))
title3d(main = "Lv 0")
subid <- currentSubscene3d()
rglwidget(elementId="rgl-pca0",width=720,height=720)
```

```{r cc-reg3d, eval=FALSE, echo=FALSE}
(pcaOt <- pca0[exType == 'other', 1:3])

names(pcaOt) <- letters[24:26]

lm.fit3d <- lm(z ~ 1 + x + y, data = pcaOt)
lm.fit3d <- lm(z ~ 0 + x + y, data = pcaOt)
Z <- predict(lm.fit3d)

(coefs <- coef(lm.fit3d))

rgl::plot3d(pcaOt[,1],pcaOt[,2],pcaOt[,3],type='s',col='blue', size=1,
            xlab = "PC1", ylab = "PC2", zlab = "PC3")
rgl::plot3d(pca0[,1],pca0[,2],pca0[,3],type='s',col=ccol3[ford], size=1,
            xlab = "PC1", ylab = "PC2", zlab = "PC3")

a <- coefs["x"]
b <- coefs["y"]
c <- 1
d <- coefs["(Intercept)"]
nv <- as.numeric(c(-a,-b,1))

D4 <- data.frame(pcaOt[,1:2], z = Z)
rgl.clear()
plot3d(D4, col='red', type = 's', size = 1, xlim= c(-5,5), ylim=c(-5,5), zlim=c(-5,5))
plot3d(pcaOt, type = 's', add=TRUE, size = 1, col = 'green')

## z = ax + by + d 
## => -ax - by + 1z - d = 0
#planes3d(-a,-b,1,0, col='red', alpha = 0.25)
#planes3d(1,1,1,0, col='red', alpha = 0.25)



r0 <- apply(pcaOt, 2, mean)
tmp <- scale(pcaOt,center=TRUE, scale=FALSE)


abclines3d(r0, a = svd(tmp)$v[,1])

(pcaOt <- pca0[exType == 'other', 1:3])
(pcaEx <- pca0[exType == 'ex', 1:3][-11,])
(pcaIn <- pca0[exType == 'in', 1:3][-6,])

colnames(pcaEx) <-
colnames(pcaIn) <-
colnames(pcaOt) <- c('x', 'y', 'z')

ex0 <- apply(pcaEx,2, mean)
in0 <- apply(pcaIn,2, mean)
ot0 <- apply(pcaOt,2, mean)
r0 <- apply(pca0, 2, mean)

vEx <- svd(scale(pcaEx,center=FALSE,scale=FALSE))$v[,1]
vIn <- svd(scale(pcaIn,center=FALSE,scale=FALSE))$v[,1]
vOt <- svd(scale(pcaOt,center=FALSE,scale=FALSE))$v[,1]

rgl.clear()

rgl::plot3d(pca0[,1],pca0[,2],pca0[,3],type='s',col=ccol3[ford], size=1,
            xlab = "PC1", ylab = "PC2", zlab = "PC3")
abclines3d(c(0,0,0), a=diag(3))

lmEx <- lm(z ~ 0 + x + y, data = pcaEx)
lmIn <- lm(z ~ 0 + x + y, data = pcaIn)
lmOt <- lm(z ~ 0 + x + y, data = pcaOt)

cEx <- coef(lmEx)
cIn <- coef(lmIn)
cOt <- coef(lmOt)

planes3d(a = -cEx[1], b = -cEx[2], c = 1, d = ex0[3], col='green', alpha=0.25)
planes3d(a = -cIn[1], b = -cIn[2], c = 1, d = in0[3], col='red', alpha=0.25)
planes3d(a = -cOt[1], b = -cOt[2], c = 1, d = ot0[3], col='blue', alpha=0.25)


abclines3d(c(0,0,0), a = vEx, col='green',size = 3)
abclines3d(c(0,0,0), a = vIn, col='red')
abclines3d(c(0,0,0), a = vOt, col='blue')

rgl::plot3d(pca0[,1],pca0[,2],pca0[,3],type='s',col=ccol3[ford], size=1,
            xlab = "PC1", ylab = "PC2", zlab = "PC3")
abclines3d(ex0, a = vEx, col='green',size = 3)
abclines3d(in0, a = vIn, col='red')
abclines3d(ot0, a = vOt, col='blue')


```


### LDA on Lv 0

Using LDA with re-substitution to create a voronoi diagram for Lv 0.
```{r cc-lda0, h = 7, w = 7, fig.cap=fig$cap("lda0", "LDA for 3 classes on Lv 0" )}
tr <- factor(exType, ordered = FALSE)
lda.fit0 <- lda(tr ~ ., data = pca0[, 1:10])
lda.pred0 <- predict(lda.fit0)

titlesvor <- paste("LDA decision boundaries for", paste0("F", 0:5))
voronoidf <- data.frame(lda.fit0$means)
#voronoidf <- data.frame(x=lda.fit$means[,1],y=lda.fit$means[,2])

#This creates the voronoi line segments

plot(pca0[,1:2], col=ccol3[ford], pch=20, cex=1.5, xlim =
     c(min(pca0[,1:2]) -3, 5))
text(pca0[,1:2], labels=rownames(pca0), 
     pos=ifelse(exType %in% c('ex', 'in'), 2, 4), 
     col=ccol3[ford], cex=1.2)

deldir(x = voronoidf[,1],y = voronoidf[,2], 
       rw = c(-15,15,-15,15), 
       plotit=TRUE, add=TRUE, wl='te')
text(voronoidf[,1:2], labels=rownames(voronoidf), cex=1.25, pos=2)
```


## Mean-difference Explorations

Staring from the correlation matrix `cmatfs` we compute the class means;
$v_1 = column\_mean($Excitatory$)$, 
$v_2 = column\_mean($Inhibitory$)$,
$v_3 = column\_mean($Other$)$.

We then compute $r_1 = v_1 - v_2$ and $r_2 = v_1 - v_3$.

We then multiply $[cmatfs] \cdot [r_1 |\, r_2]$.
The transformed points are then plotted below.

```{r cc-msp}
tmp <- cmatfs; diag(tmp) <- 0
exMat <- tmp[exType == 'ex',] -> g1
inMat <- tmp[exType == 'in',] -> g2
otMat <- tmp[exType == 'other',] -> g3

mEx <- apply(exMat, 2, mean) -> v1   
mIn <- apply(inMat, 2, mean) -> v2 
mOt <- apply(otMat, 2, mean) -> v3 

mXI <- mEx - mIn
mXO <- mEx - mOt

rotM <- data.frame(XI = mXI, XO = mXO) -> r1r2
colnames(r1r2) <- c("r1", "r2")

mdm2 <- data.frame(cmatfs %*% as.matrix(rotM))
mdm1 <- data.frame(XI = mdm2[,1], row.names=rownames(mdm2))
newFord <- order(mdm2$XI)
```

### Within Group Heatmaps

```{r cc-heatG123,eval=TRUE, fig.show='hold',w=6,h=6,fig.cap=fig$cap("heatg123", "Heatmaps of groups Ex, In, Ot" )}
heatmap.2(as.matrix(g1[, newFord]),dendrogram='none',
          Colv=NA,trace="none", 
          col=mycol,colCol=ccol3[ford][newFord],
          cexRow=0.8, keysize=1.25,
          symkey=FALSE,symbreaks=FALSE,
          scale="none", srtCol=90,
          main="Excitatory") 

heatmap.2(as.matrix(g2[, newFord]),dendrogram='none',
          Colv=NA,trace="none", 
          col=mycol,colCol=ccol3[ford][newFord],
          cexRow=0.8, keysize=1.25,
          symkey=FALSE,symbreaks=FALSE,
          scale="none", srtCol=90,
          main="Inhibatory") 

heatmap.2(as.matrix(g3[, newFord]),dendrogram='none',
          Colv=NA,trace="none", 
          col=mycol,colCol=ccol3[ford][newFord],
          cexRow=0.8, keysize=1.25,
          symkey=FALSE,symbreaks=FALSE,
          scale="none", srtCol=90,
          main="Other") 
rm(tmp)
```

### Group Mean Heatmaps

```{r cc-heatV123,eval=TRUE, fig.show='hold',w=6,h=4,fig.cap=fig$cap("heatV123", "Heatmaps of v1, v2, v3" )}
tmp <- t(cbind(v1, v1))[, newFord]
heatmap.2(as.matrix(tmp),dendrogram='none',
          Colv=NA,trace="none", key= FALSE,
          col=mycol,colCol=ccol3[ford][newFord],
          cexRow=0.8, keysize=1.25,
          symkey=FALSE,symbreaks=FALSE,
          scale="none", srtCol=90, labRow='',
          main="v1 = Mean Excitatory") 

tmp <- t(cbind(v2, v2))[, newFord]
heatmap.2(as.matrix(tmp),dendrogram='none',
          Colv=NA,trace="none", key = FALSE,
          col=mycol,colCol=ccol3[ford][newFord],
          cexRow=0.8, keysize=1.25,
          symkey=FALSE,symbreaks=FALSE,
          scale="none", srtCol=90, labRow='',
          main="v2 = Mean Inhibitory") 

tmp <- t(cbind(v3, v3))[, newFord]
heatmap.2(as.matrix(tmp),dendrogram='none',
          Colv=NA,trace="none", key = FALSE,
          col=mycol,colCol=ccol3[ford][newFord],
          cexRow=0.8, keysize=1.25,
          symkey=FALSE,symbreaks=FALSE,
          scale="none", srtCol=90, labRow='',
          main="v3 = Mean Other") 
rm(tmp)
```

### Mean Difference Vectors Heatmap

```{r cc-heatR1R2, eval=TRUE,w=6,h=6,fig.cap=fig$cap("heatR1R2", "Heatmap of the mean difference vectors r1 and r2." )}
heatmap.2(as.matrix(t(r1r2)[, newFord]),dendrogram='none',
          Colv=NA,trace="none", 
          col=mycol,colCol=ccol3[ford][newFord],
          cexRow=0.8, keysize=1.25,
          symkey=FALSE,symbreaks=FALSE,
          scale="none", srtCol=90,
          main="Mean difference vectors") 
```

### Mean-difference Marker Projections

```{r p-mdp2, results='hold', w=7,h=7,fig.cap=fig$cap("mdp2", "Mean-difference projections 2d")}
plot(mdm2, col = ccol3[ford], pch = 20, cex=1)
abline(h = 0, v = 0)
text(mdm2, labels=rownames(mdm2), col=ccol3[ford], cex = 0.75, pos=1)


lm.fit <- lm(mdm2[,2] ~ mdm2[,1])
rL <- list(a = lm.fit$coefficients[1],
          b = lm.fit$coefficients[2])
L1 <- function(x){ rL$a + rL$b * x }

summary(lm.fit)

abline(rL$a,rL$b, col = 'darkorange', lty=2)
text(-0.1,L1(0.1) , labels=expression(L[1]), col='darkorange', cex = 1) 

legend(list(x= -1.5,y=1.6), legend = c('regression line'), col = 'darkorange', lty=2)
title("Mean-difference projections of markers")
```


The following is the projection onto $r_1$. 

```{r p-mdp1, w=7,h=3,fig.cap=fig$cap("mdp1", "Mean-difference projections 1d")}
set.seed(2^8)
print(ggplot(mdm1, aes(x = XI, y = 0, label=rownames(mdm1))) +
  geom_point(color = ccol3[ford]) + 
  geom_text(color = ccol3[ford], check_overlap=FALSE,
            position='jitter'))
```

We will now project the individual synapse points onto the  
line $L_1$ by first projecting into the mean-difference
space then projecting onto $L_1$ followed by a rotation into
$\mathbb{R}$.

```{r cc-synProj21, eval=FALSE}
mfs <- as.matrix(fs)
r <- as.matrix(mdm1)
projR1 <- mfs %*% r

plot(density(projR1))

if(FALSE){
require(ggvis)
data.frame(projR1) %>% ggvis(x = ~XI) %>%
    layer_densities(
      adjust = input_slider(.1, 2, value = 1, step = .1, label = "Bandwidth adjustment"),
      kernel = input_select(
        c("Gaussian" = "gaussian",
          "Epanechnikov" = "epanechnikov",
          "Rectangular" = "rectangular",
          "Triangular" = "triangular",
          "Biweight" = "biweight",
          "Cosine" = "cosine",
          "Optcosine" = "optcosine"),
        label = "Kernel")
    )
}
```

```{r cc-BIC, eval=TRUE, echo=TRUE}
#bic on mdm1
    n <- nrow(mdm1)
    d <- ncol(mdm1)
    G <- 3
    emDat <- cbind(mdm1, type=as.numeric(exType))
    emEst <- me(modelName = 'V', data=as.matrix(emDat[, -2]), unmap(emDat[,2]))

    #names(emEst)
    #args(bic)
    #do.call('bic', emEst)

    B <- foreach(i = 1:25, .combine='c') %do% {
      bic(modelName="V", loglik=emEst$loglik, n=n, d=d, G=i)
    }
    plot(B, type = 'b')
```

```{r cc-svdRSV,eval=TRUE,w=6,h=6,fig.cap=fig$cap("svdRSV", "Heatmap of 1:2 right singular vectors of 'fs' data ordered by projection on L_1." )}
svdfs <- svd(as.matrix(fs)[,newFord])

rightSV <- data.frame(t(svdfs$v[,1:2]))
colnames(rightSV) <- rownames(mdm2[order(mdm2$XI),])

heatmap.2(as.matrix(rightSV),dendrogram='row',Colv=NA,trace="none", col=mycol,colCol=ccol3[ford][newFord],cexRow=0.8, keysize=1.25,symkey=FALSE,symbreaks=FALSE,scale="none", srtCol=90,main="r. singular vectors of 'fs' data") 
```

# Level 1: K-means++ for $K=2$. 
We run a Hierachical K-means++ for $K=2$ on the `fs` data with 4 levels.

```{r cc-kmpp,eval=TRUE,echo=TRUE}
set.seed(2^13)
L <- bhkmpp(dat,blevels=4)
```

## Heat maps (Lv 1):

```{r cc_agg, eval=TRUE}
## Formatting data for heatmap
aggp <- aggregate(dat,by=list(lab=L[[1]]),FUN=mean)
aggp <- as.matrix(aggp[,-1])[, ford]
rownames(aggp) <- clusterFraction(L[[1]])
```

The following are heatmaps generated from clustering 
via K-means++ (at level 1)

```{r cc-kmpp-heatmapLV1,eval=FALSE,echo=FALSE,w=6,h=6,fig.cap=fig$cap("heat1", "Heatmap of the cluster means vs channels. Rows and columns are rearranged according to hclust.")}
heatmap.2(as.matrix(aggp), trace="none",col=mycol,colCol=ccol,cexRow=0.8, keysize=1,symkey=FALSE,symbreaks=FALSE,scale="none", srtCol=90,main="Heatmap of `fs` data") # 
```

```{r cc-kmpp-heatmapSortedLV1,eval=TRUE,w=6,h=6,fig.cap=fig$cap("heatfs1", "Heatmap of the cluster means vs channels. Rows and columns are rearranged according to synapse type.")}
heatmap.2(as.matrix(aggp),dendrogram='row',Colv=NA,trace="none", col=mycol,colCol=ccol[ford],cexRow=0.8, keysize=1.25,symkey=FALSE,symbreaks=FALSE,scale="none", srtCol=90,main="Heatmap of `fs` data.") 
```

Percentage of data within cluster is presented on the right side of the heatmap.


## Jittered scatter plot: Lv 1


```{r ggJitter1}
ggCol <- brewer.pal(4,"Set1")[order(table(L[[1]]))]

cf1 <- data.frame(cf = clusterFraction(L[[1]]))

ggJ1 <- 
  ggplot(data = ggJdat, aes(x = ind, y = values, 
                         color = as.factor(lv1))) +
  scale_color_manual(values=ggCol, name="Cluster") + 
  geom_point(alpha=0.25, position=position_jitterdodge()) + 
  geom_boxplot(alpha =0.35, outlier.color = 'NA') + 
  annotate("text", x = levels(ggJdat$ind)[c(2,20)], y = 1.15*max(ggJdat$values), 
           label= cf1[1:2,]) + 
  theme(axis.title.x = element_blank()) + 
  theme(axis.text.x = element_text(color = ccol[ford], 
                                   angle=45,
                                   vjust = 0.5))
```

```{r jitter1, h=6, w=14, fig.cap=fig$cap("jitter1","Scatter Plot Level 1 ")}
print(ggJ1)
```

## Within cluster correlations (Lv 1)

```{r cc-wcc1, h=8,w=8,fig.cap=fig$cap("corkp1","Within cluster correlations, clock-wise from top left, Cluster 1, Cluster 2, difference C1 - C2")}
corkp1 <- cor(dat[L[[1]] == 1,])[ford, ford]
corkp2 <- cor(dat[L[[1]] == 2,])[ford, ford]
difCor12 <- (corkp1 - corkp2)

layout(matrix(c(1,2,3,3), 2, 2, byrow=TRUE))
corrplot(corkp1,method="color",tl.col=ccol[ford], tl.cex=0.8, mar=c(0,0,3,0))
title("Cluster 1")
corrplot(corkp2,method="color",tl.col=ccol[ford], tl.cex=0.8, mar=c(0,0,3,0))
title("Cluster 2")
corrplot(difCor12,is.corr=FALSE,method="color",
         tl.col=ccol[ford], tl.cex=0.8, 
         mar=c(0,0,3,0),
         col=colorRampPalette(c("#998ec3","white","darkorange"))(50))
title("Difference(1,2)")
```

Notice that the non-synaptic markers change very little between
clusters.  Also note that the correlations between (`gad, VGAT, PV,
Gephyr`) and `VGlut1` at both times change significantly between
clusters.

Next we compute PCA for the within cluster correlation matrices and
embed in 3d.

### PCA of the within cluster correlation matrices at level 1.
```{r cc-pca1,h=7,w=7,fig.cap=fig$cap("pairs1","PCA of untransformed correlation matrix")}
pcaL <- lapply(list(corkp1, corkp2), prcomp, center=TRUE, scale=TRUE)
elB <- lapply(pcaL, function(x) {getElbows(x$sdev, plot=FALSE)})
```

#### PCA C1 Lv1
```{r cc_3dCorr1,echo=TRUE, eval=TRUE, fig.cap=fig$cap("rgl1","PCA 1-3 on untransformed correlation matrix")}
pca <- pcaL[[1]]$x
rgl::plot3d(pca[,1],pca[,2],pca[,3],type='s',col=ccol3[ford], size=1,
            xlab = "PC1", ylab = "PC2", zlab = "PC3")
rgl::rgl.texts(pca[,1],pca[,2],pca[,3],abbreviate(rownames(pca)), col=ccol3[ford], adj=c(0,2))
title3d(main = "Cluster 1: Lv 1")
subid <- currentSubscene3d()
rglwidget(elementId="rgl-pca1",width=720,height=720)
```

#### PCA C1 Lv1
```{r cc_3dCorr2,echo=TRUE, eval=TRUE, fig.cap=fig$cap("rgl2","PCA 1-3 on untransformed correlation matrix")}
pca <- pcaL[[2]]$x
rgl::plot3d(pca[,1],pca[,2],pca[,3],type='s',col=ccol3[ford], size=1,
            xlab = "PC1", ylab = "PC2", zlab = "PC3")
rgl::rgl.texts(pca[,1],pca[,2],pca[,3],abbreviate(rownames(pca)), col=ccol3[ford], adj=c(0,2))
title3d(main = "Cluster 2: Lv 1")
subid <- currentSubscene3d()
rglwidget(elementId="rgl-pca2",width=720,height=720)
```

### LDA on Lv1

```{r cc-voronoi1Lv1,h=8,w=12,fig.cap=fig$cap("voronoiLv1","Voronoi diagrams on class means from LDA on PCA of untransformed correlation matrices")}
lda.fit <- 
  lapply(list(pcaL[[1]]$x[,1:elB[[1]][2]],
              pcaL[[2]]$x[,1:elB[[2]][2]]),
                function(y) {
                  lda(tr ~ ., data = as.data.frame(y))
                })

titlesvor <- paste("LDA decision boundaries for", paste0("C", 1:2))
voronoidf <- lapply(lapply(lda.fit, '[[', 3), data.frame)

#This creates the voronoi line segments

par(mfrow = c(1,2))
for(i in 1:length(pcaL)){
  plot(pcaL[[i]]$x[,1:2], col=ccol3[ford], pch=20, cex=1.5)
  title(titlesvor[i])
  text(pcaL[[i]]$x[,1:2], labels=rownames(pcaL[[i]]$x), 
       pos=ifelse(pcaL[[i]]$x[,1]<max(pcaL[[i]]$x[,1] -0.5),4,2), 
       col=ccol3[ford], cex=1.2)

  deldir(x = voronoidf[[i]][,1],y = voronoidf[[i]][,2], rw = c(-15,15,-15,15), 
       plotit=TRUE, add=TRUE, wl='te')
  text(voronoidf[[i]], labels=rownames(voronoidf[[i]]), cex=1.5, pos=1)
}
```


## Clusters and Spatial Location (Lv 1)
We show a 3d scatter plot of the data according to spatial location
colored according to cluster labels as determined by k-means++.

The reader should notice from this figure that there does not 
seem to be any correlation between the spatial information and cluster labels.



```{r cc-kpp3dLV1}
set.seed(2^12)
s1 <- sample(dim(loc)[1],5e4)

locs1 <- loc[s1,]
locs1$cluster <- L[[1]][s1]

plot3d(locs1$V1,locs1$V2,locs1$V3,
       col=brewer.pal(4,"Set1")[order(table(L[[1]]))][locs1$cluster],
       alpha=0.75,
       xlab='x', 
       ylab='y', 
       zlab='z')

subid <- currentSubscene3d()
rglwidget(elementId="plot3dLocations", height=720, width=720)
```


# Level 2: K-means++ for $K=2$.


## Heat maps (Lv 2):

```{r cc-agg2, eval=TRUE}
## Formatting data for heatmap
aggp2 <- aggregate(dat,by=list(lab=L[[2]]),FUN=function(x){mean(x)}) 
aggp2 <- as.matrix(aggp2[,-1])[, ford]
rownames(aggp2) <- clusterFraction(L[[2]])
```

The following are heatmaps generated from clustering 
via K-means++

```{r cc-kmpp-heatmapLV2,eval=FALSE,echo=FALSE,w=6,h=6,fig.cap=fig$cap("heat1", "Heatmap of the cluster means vs channels. Rows and columns are rearranged according to hclust.")}
heatmap.2(as.matrix(aggp2), trace="none",col=mycol,colCol=ccol,cexRow=0.8, keysize=1,symkey=FALSE,symbreaks=FALSE,scale="none", srtCol=90,main="Heatmap of `fs` data") # 
```

```{r cc-kmpp-heatmapSortedLV2,eval=TRUE,w=6,h=6,fig.cap=fig$cap("heatfs1", "Heatmap of the cluster means vs channels. Rows and columns are rearranged according to synapse type.")}
heatmap.2(as.matrix(aggp2),dendrogram='row',Colv=NA,trace="none", col=mycol,colCol=ccol[ford],cexRow=0.8, keysize=1.25,symkey=FALSE,symbreaks=FALSE,scale="none", srtCol=90,main="Heatmap of `fs` data.") 
```

Percentage of data within cluster is presented on the right side of the heatmap.


## Jittered scatter plot: Lv 2


```{r ggJitter2}
ggCol <- brewer.pal(8,"Set1")[order(table(L[[2]]))]
cf2 <- data.frame(cf = clusterFraction(L[[2]]))

ggJ2 <- 
  ggplot(data = ggJdat, aes(x = ind, y = values, 
                         color = as.factor(lv2))) +
  scale_color_manual(values=ggCol, name="Cluster") + 
  geom_point(alpha=0.25, position=position_jitterdodge()) + 
  geom_boxplot(alpha =0.35, outlier.color = 'NA') + 
  annotate("text", x = levels(ggJdat$ind)[c(2,8,14,20)], y = 1.15*max(ggJdat$values), 
           label= cf2[1:4,]) + 
  theme(axis.title.x = element_blank()) + 
  theme(axis.text.x = element_text(color = ccol[ford], 
                                   angle=45,
                                   vjust = 0.5))
```

```{r jitter2, h=6, w=14, fig.cap=fig$cap("jitter2","Scatter Plot Level 2 ")}
print(ggJ2)
```

The fraction of data points within each cluster are given at the top of
the plot window. 


## Within cluster correlations (Lv 2)

```{r cc-wccLV2,h=12,w=7,fig.cap=fig$cap("corkp2","Within cluster correlations for level 2. (c11, c12, c21, c22) with differences")}
corLV2 <- lapply(c(1:4),function(x){cor(dat[L[[2]] == x,])[ford, ford]})

difCor1112 <- ((corLV2[[1]] - corLV2[[2]]))
difCor2122 <- ((corLV2[[3]] - corLV2[[4]]))

layout(matrix(c(1,2,3,3,4,5,6,6), 4, 2, byrow=TRUE))
corrplot(corLV2[[1]],method="color",tl.col=ccol[ford], tl.cex=0.8, 
         mar=c(0,0,3,0))
title("Cluster 1")
corrplot(corLV2[[2]],method="color",tl.col=ccol[ford], tl.cex=0.8, 
         mar=c(0,0,3,0))
title("Cluster 2")
corrplot(difCor1112, method="color", tl.col=ccol[ford], 
         tl.cex=0.8,
         mar = c(0,0,3,0),
         cl.lim = c(min(difCor1112,difCor2122),max(difCor1112,difCor2122)),
         col=colorRampPalette(c("#998ec3", 
                                "white",
                                "darkorange"))(100))
title("Difference(1,2)")
corrplot(corLV2[[3]],method="color",tl.col=ccol[ford], tl.cex=0.8, 
         mar=c(0,0,3,0))
title("Cluster 3")
corrplot(corLV2[[4]],method="color",tl.col=ccol[ford], tl.cex=0.8, 
         mar=c(0,0,3,0))
title("Cluster 4")
corrplot(difCor2122, method="color", tl.col=ccol[ford], 
         tl.cex=0.8,
         mar=c(0,0,3,0),
         cl.lim = c(min(difCor1112,difCor2122),max(difCor1112,difCor2122)),
         col=colorRampPalette(c("#998ec3", 
                                "white",
                                "darkorange"))(100))
title("Difference(3,4)")
```


## PCA of the within cluster correlation matrices at level 2.
```{r cc-pca1lv2,h=7,w=7,fig.cap=fig$cap("pairsLV2","PCA of untransformed correlation matrix")}
pcaL <- lapply(corLV2, prcomp, center=TRUE, scale=TRUE)
elB <- lapply(pcaL, function(x) {getElbows(x$sdev, plot=FALSE)})
pcaLel2 <- mapply(function(x,y){ x$x[,1:y[2]] }, pcaL, elB)
```

## LDA on Lv 2

```{r cc-voronoiLv2,h=10,w=12,fig.cap=fig$cap("voronoiLv2","Voronoi diagrams on class means from LDA on PCA of untransformed correlation matrices")}
lda.fit <- 
  lapply(pcaLel2, 
         function(y) {
                  lda(tr ~ ., data = as.data.frame(y))
                })

titlesvor <- paste("LDA decision boundaries for", paste0("C", 1:4))
voronoidf <- lapply(lapply(lda.fit, '[[', 3), data.frame)

#This creates the voronoi line segments

par(mfrow = c(2,2))
for(i in 1:length(pcaL)){
  plot(pcaL[[i]]$x[,1:2], col=ccol3[ford], pch=20, cex=1.5)
  title(titlesvor[i])
  text(pcaL[[i]]$x[,1:2], labels=rownames(pcaL[[i]]$x), 
       pos=ifelse(pcaL[[i]]$x[,1]<max(pcaL[[i]]$x[,1] -0.5),4,2), 
       col=ccol3[ford], cex=1.2)

  deldir(x = voronoidf[[i]][,1],y = voronoidf[[i]][,2], rw = c(-15,15,-15,15), 
       plotit=TRUE, add=TRUE, wl='te')
  text(voronoidf[[i]], labels=rownames(voronoidf[[i]]), cex=1.5, pos=1)
}
```





## Clusters and Spatial Location (Lv 2)
Using the location data and the results of K-means++ we show a 3d scatter
plot colored according to cluster.

```{r cc-kpp3d2}
set.seed(2^12)
s1 <- sample(dim(loc)[1],5e4)

locs2 <- loc[s1,]
locs2$cluster <- L[[2]][s1]

YlOrBr <- c("#FFFFD4", "#FED98E", "#FE9929", "#D95F0E", "#993404")
col.pal <- colorRampPalette(YlOrBr)

plot3d(locs2$V1,locs2$V2,locs2$V3,
       #col=colorpanel(8,"brown","blue")[order(table(L[[2]]))][locs2$cluster],
       col=col.pal(8)[-seq(1,8,2)][order(table(L[[2]]))][locs2$cluster],
       alpha=0.75,
       xlab='x', 
       ylab='y', 
       zlab='z'
       )

subid <- currentSubscene3d()
rglwidget(elementId="plot3dLocationsLV2", height=720, width=720)
```




# Level 3: K-means++ for $K=2$.


## Heat maps (Lv 3):

```{r cc-agg3, eval=TRUE}
## Formatting data for heatmap
aggp3 <- aggregate(dat,by=list(lab=L[[3]]),FUN=function(x){mean(x)})
aggp3 <- as.matrix(aggp3[,-1])[, ford]
rownames(aggp3) <- clusterFraction(L[[3]])
```

The following are heatmaps generated from clustering 
via K-means++

```{r cc-kmpp-heatmapLV3,eval=FALSE,echo=FALSE,w=6,h=6,fig.cap=fig$cap("heat3", "Heatmap of the cluster means vs channels. Rows and columns are rearranged according to hclust.")}
heatmap.2(as.matrix(aggp3), trace="none",col=mycol,colCol=ccol,cexRow=0.8, keysize=1,symkey=FALSE,symbreaks=FALSE,scale="none", srtCol=90,main="Heatmap of `fs` data") # 
```

```{r cc-kmpp-heatmapSortedLV3,eval=TRUE,w=6,h=6,fig.cap=fig$cap("heatfs1", "Heatmap of the cluster means vs channels. Rows and columns are rearranged according to synapse type.")}
heatmap.2(as.matrix(aggp3),dendrogram='row',Colv=NA,trace="none", col=mycol,colCol=ccol[ford],cexRow=0.8, keysize=1.25,symkey=FALSE,symbreaks=FALSE,scale="none", srtCol=90,main="Heatmap of `fs` data.") 
```

Percentage of data within cluster is presented on the right side of the heatmap.


## Jittered scatter plot: Lv 3


```{r ggJitter3}
ggCol <- brewer.pal(8,"Set1")[order(table(L[[3]]))]
cf3 <- data.frame(cf = clusterFraction(L[[3]]))

ggJ3 <- 
  ggplot(data = ggJdat, aes(x = ind, y = values, 
                         color = as.factor(lv3))) +
  scale_color_manual(values=ggCol, name="Cluster") + 
  geom_point(alpha=0.25, position=position_jitterdodge()) + 
  geom_boxplot(alpha =0.35, outlier.color = 'NA') + 
  annotate("text", x = levels(ggJdat$ind)[seq(2,22,length=8)], y = 1.05*max(ggJdat$values), 
           label= cf3[1:8,]) + 
  #geom_jitter(width=2) + 
  theme(axis.title.x = element_blank()) + 
  theme(axis.text.x = element_text(color = ccol[ford], 
                                   angle=45,
                                   vjust = 0.5))
```

```{r jitter3, h=6, w=17, fig.cap=fig$cap("jitter3","Scatter Plot Level 3 ")}
print(ggJ3)
```

## Within cluster correlations (Lv 3)

```{r cc-wcc3,h=17,w=5,fig.cap=fig$cap("corkp3","Within cluster correlations for level 3. (c111, c112, c121, c122, c211, c212, c221, c222)")}
corLV3 <- lapply(c(1:8),function(x){cor(dat[L[[3]] == x,])[ford, ford]})

difCor1 <- (corLV3[[1]] - corLV3[[2]])
difCor2 <- (corLV3[[3]] - corLV3[[4]])
difCor3 <- (corLV3[[5]] - corLV3[[6]])
difCor4 <- (corLV3[[7]] - corLV3[[8]])
M <- max(difCor1, difCor2, difCor3, difCor4)
m <- min(difCor1, difCor2, difCor3, difCor4)

layout(matrix(c(1, 2, 3, 3,
                4, 5, 6, 6, 
                7, 8, 9, 9,
                10, 11, 12, 12), 8,2, byrow=TRUE))

corrplot(corLV3[[1]],method="color",tl.col=ccol[ford], tl.cex=0.8,
         mar=c(0,0,3,0))
title('Cluster 1')
corrplot(corLV3[[2]],method="color",tl.col=ccol[ford], tl.cex=0.8,
         mar=c(0,0,3,0))
title('Cluster 2')
corrplot(difCor1,method="color",tl.col=ccol[ford], tl.cex=0.8,
         cl.lim=c(m,M), 
         mar=c(0,0,3,0),
         col=colorRampPalette(c("#998ec3",
                                "white",
                                "darkorange"))(50))
title('Difference(1,2)')
corrplot(corLV3[[3]],method="color",tl.col=ccol[ford], tl.cex=0.8, 
         mar=c(0,0,3,0))
title('Cluster 3')
corrplot(corLV3[[4]],method="color",tl.col=ccol[ford], tl.cex=0.8, 
         mar=c(0,0,3,0))
title('Cluster 4')
corrplot(difCor2,method="color",tl.col=ccol[ford], tl.cex=0.8, 
         cl.lim= c(m,M),
         mar=c(0,0,3,0),
         col=colorRampPalette(c("#998ec3",
                                "white",
                                "darkorange"))(50))
title('Difference(3,4)')
corrplot(corLV3[[5]],method="color",tl.col=ccol[ford], tl.cex=0.8, 
         mar=c(0,0,3,0))
title('Cluster 5')
corrplot(corLV3[[6]],method="color",tl.col=ccol[ford], tl.cex=0.8, 
         mar=c(0,0,3,0))
title('Cluster 6')
corrplot(difCor3,method="color",tl.col=ccol[ford], tl.cex=0.8,
         cl.lim= c(m,M),
         mar=c(0,0,3,0),
         col=colorRampPalette(c("#998ec3",
                                "white",
                                "darkorange"))(50))
title('Difference(5,6)')
corrplot(corLV3[[7]],method="color",tl.col=ccol[ford], tl.cex=0.8, 
         mar=c(0,0,3,0))
title('Cluster 7')
corrplot(corLV3[[8]],method="color",tl.col=ccol[ford], tl.cex=0.8, 
         mar=c(0,0,3,0))
title('Cluster 8')
corrplot(difCor4,method="color",tl.col=ccol[ford], tl.cex=0.8,
         cl.lim= c(m,M),
         mar=c(0,0,3,0),
         col=colorRampPalette(c("#998ec3",
                                "white",
                                "darkorange"))(50))
title('Difference(7,8)')
```


## PCA of the within cluster correlation matrices at level 3.
```{r cc-pca1lv3}
pcaL <- lapply(corLV3, prcomp, center=TRUE, scale=TRUE)
elB <- lapply(pcaL, function(x) {getElbows(x$sdev, plot=FALSE)})
pcaLel3 <- mapply(function(x,y){ x$x[,1:y[2]] }, pcaL, elB)
```

## LDA on Lv 3

```{r cc-voronoiLv3,h=18,w=12,fig.cap=fig$cap("voronoiLv3","Voronoi diagrams on class means from LDA on PCA of untransformed correlation matrices")}
lda.fit <- 
  lapply(pcaLel3, 
         function(y) {
                  lda(tr ~ ., data = as.data.frame(y))
                })

titlesvor <- paste("LDA decision boundaries for", paste0("C", 1:8))
voronoidf <- lapply(lapply(lda.fit, '[[', 3), data.frame)

#This creates the voronoi line segments

par(mfrow = c(4,2))
for(i in 1:length(pcaL)){
  plot(pcaL[[i]]$x[,1:2], col=ccol3[ford], pch=20, cex=1.5)
  title(titlesvor[i])
  text(pcaL[[i]]$x[,1:2], labels=rownames(pcaL[[i]]$x), 
       pos=ifelse(pcaL[[i]]$x[,1]<max(pcaL[[i]]$x[,1] -0.5),4,2), 
       col=ccol3[ford], cex=1.2)

  deldir(x = voronoidf[[i]][,1],y = voronoidf[[i]][,2], rw = c(-15,15,-15,15), 
       plotit=TRUE, add=TRUE, wl='te')
  text(voronoidf[[i]], labels=rownames(voronoidf[[i]]), cex=1.5, pos=1)
}
```


## Clusters and Spatial Location (Lv 3)
Using the location data and the results of K-means++ we show a 3d scatter
plot colored according to cluster.

```{r cc-kppLV3}
set.seed(2^12)
s1 <- sample(dim(loc)[1],5e4)

locs3 <- loc[s1,]
locs3$cluster <- L[[3]][s1]

plot3d(locs3$V1,locs3$V2,locs3$V3,
       col=col.pal(16)[-seq(1,8,2)][order(table(L[[3]]))][locs3$cluster],
       alpha=0.65,
       xlab='x', 
       ylab='y', 
       zlab='z'
       )

subid <- currentSubscene3d()
rglwidget(elementId="plot3dLocationsLV3", height=720, width=720)
```


<footer>
<p> [Back to top][Definition of Markers]</p>
</footer>
