---
title: "RORB Data: pseudo-centroids from annotation polygons using Gaussian weighted average over 11x11x11 cubes  ≈ 0.61 microns^3 followed by inhibitory/excitatory analysis."
date: '`r Sys.Date()`'
output:
  html_document:
    fig_caption: yes
    fig_height: 5
    fig_width: 5
    highlight: pygments
    number_sections: yes
    theme: cerulean
    toc: yes
    toc_depth: 2
---
```{r knitOPTS, include=FALSE}
knitr::opts_chunk$set(cache = FALSE, dev = "png")
```

```{r render, eval=FALSE, echo=FALSE}
require(rmarkdown)
require(knitr)
rmarkdown::render("rorb_meda_plots_inex.Rmd")
system("open rorb_meda_plots_inex.html -a /Applications/Chrome.app")
```

```{r setup,include=FALSE,results='asis',message=FALSE,warning=FALSE, echo = FALSE}
# Library calls here.
require(rmarkdown)
require(knitr)
require(doMC)
require(foreach)
require(scales)
require(kernlab)
require(dimRed)
require(car)
registerDoMC(4)
suppressMessages(require(meda))
```

```{r, eval = TRUE, echo = FALSE, include = FALSE}
dat <- read.csv("rorb_gaussianAvg_at.csv")
loc <- read.csv("rorb_gaussianAvg_at_orderLocations.csv")
gabaID <- read.csv("rorb_gaba.csv")
truth <- gaba <- gabaID$gaba

ccol <- c('blue', 'blue', 'blue', 'red', 'red', 'red', 'green', 'black', 'green', 'green', 'black', 'green')
ind <- order(ccol)
ccol <- sort(ccol)

dat <- dat[,ind]
sdat <- as.data.frame(scale(dat, center = TRUE, scale = TRUE))
#stackMraw(sdat, as.factor(gaba), ccol = ccol, centered = TRUE, depth = 2) + ggtitle("True Clusters")
```


# In/Ex

I start with the Gaussian weighted averages over the `11 x 11 x 11 x Ch` cubes at
each synapse location.

Here I re-factor the data into a two-variable dataset: 

INH $:= \text{Z-score} \circ \sum{\text{Inhibitory Features}}$

EX $:= \text{Z-score} \circ \sum{\text{Excitatory Features}}$

Then plot the new dataset colored by the given `gaba` labels.
The ellipses superimposed on the plot correspond to the class conditional covariances (at 1,2, and 3 standard deviations) of
the components.


```{r inex, fig.width = 8, fig.height = 8, echo = FALSE}
ndat <- data.frame(
                   "inhibitory" = Reduce('+', sdat[, ccol == 'green']), 
                   "excitatory" = Reduce('+', sdat[, ccol == 'red'])
                   )

sndat <- data.frame(gaba = gaba, scale(ndat, center = TRUE, scale = TRUE))

plot(sndat[, -1], type = 'n')
points(sndat[gaba == 0, -1], col = alpha('darkblue', 0.5), pch = 20, cex = 0.5)
points(sndat[gaba == 1, -1], col = alpha('red', 0.5), pch = 20, cex = 0.5)
abline(h = 0)
abline(v = 0)


points(colMeans(sndat[gaba == 0, -1])[1],colMeans(sndat[gaba == 0, -1])[2], col = 'blue', cex = 3)
points(colMeans(sndat[gaba == 0, -1])[1],colMeans(sndat[gaba == 0, -1])[2], col = 'blue', cex = 2, pch = 3)
points(colMeans(sndat[gaba == 1, -1])[1],colMeans(sndat[gaba == 1, -1])[2], col = 'red', cex = 3)
points(colMeans(sndat[gaba == 1, -1])[1],colMeans(sndat[gaba == 1, -1])[2], col = 'red', cex = 2, pch = 3)


A <- cov(sndat[gaba == 0, -1])
B <- cov(sndat[gaba == 1, -1])

ctr <- colMeans(sndat[gaba == 0, -1])
RR <- chol(A)

angles <- seq(0, 2*pi, length.out=200)          # angles for ellipse
ell    <- 1 * cbind(cos(angles), sin(angles)) %*% RR  # ellipse scaled with factor 1
ellCtr <- sweep(ell, 2, ctr, "+")               # center ellipse to the data centroid
#plot(ellCtr, type="l", lwd=2, asp=1)            # plot ellipse
#points(ctr[1], ctr[2], pch=4, lwd=2)            # plot data centroid
ellipse(ctr, shape=A, radius=1, col=alpha("blue", 0.5), lty=2, add = TRUE, center.pch = 3)
ellipse(ctr, shape=A, radius=2, col=alpha("blue", 0.5), lty=2, add = TRUE, center.pch = 0)
ellipse(ctr, shape=A, radius=3, col=alpha("blue", 0.5), lty=2, add = TRUE, center.pch = 0)


ctr <- colMeans(sndat[gaba == 1, -1])
RR <- chol(B)


angles <- seq(0, 2*pi, length.out=200)          # angles for ellipse
ell    <- 1 * cbind(cos(angles), sin(angles)) %*% RR  # ellipse scaled with factor 1
ellCtr <- sweep(ell, 2, ctr, "+")               # center ellipse to the data centroid

ellipse(ctr, shape=B, radius=1, col=alpha("red",0.5), lty=2, add = TRUE, center.pch = 3)
ellipse(ctr, shape=B, radius=2, col=alpha("red",0.5), lty=2, add = TRUE, center.pch = 0)
ellipse(ctr, shape=B, radius=3, col=alpha("red",0.5), lty=2, add = TRUE, center.pch = 0)
```



