---
title: "RORB Data: "
date: '`r Sys.Date()`'
output:
  html_document:
    fig_caption: yes
    fig_height: 8
    fig_width: 8
    highlight: pygments
    number_sections: no
    theme: cerulean
    toc: yes
    toc_depth: 2
---
```{r knitOPTS, include=FALSE}
knitr::opts_chunk$set(cache = FALSE, dev = "png", results = 'hold')
```

```{r render, eval=FALSE, echo=FALSE}
require(rmarkdown)
require(knitr)
rmarkdown::render("verify.Rmd")
system("open verify.html -a /Applications/Chrome.app")
```

```{r setup,include=FALSE,results='asis',message=FALSE,warning=FALSE, echo = FALSE}
# Library calls here.
require(rmarkdown)
require(knitr)
require(doMC)
require(foreach)
require(scales)
require(kernlab)
require(dimRed)
require(randomForest)
require(mvtnorm)
require(class)
require(MASS)
registerDoMC(6)
suppressMessages(require(meda))
```


# Verifiy and Extend
> Verify and extend this, Jesse:
> 
> by my calculations,
> letting
>   $\hat{L}^{(R)}_d$
> denote the resubstitution error rate of the quadratic Bayes plug-in classifier using $d$ dimensions
> when $f_0 = MVN([0,0],I)$ and $f_1 = MVN([1,0],I)$,
> conditioning on $n_0=n_1=n/2$
> with $n=20$
> yields
> $$\mathbb{P}[ \hat{L}^{(R)}_2 > \hat{L}^{(R)}_1 ] \approx 0.18$$
> and
> $$\mathbb{P}[ \hat{L}^{(R)}_2 = \hat{L}^{(R)}_1 ] \approx 0.29$$
> that is,
> the probability that the second (useless) feature *degrades* performance,
> even when testing on the training set, is substantial.

---

## Emprical results:

### 500 Monticarlo replicates with $f_0 = MVN([0,0,0,0], I)$ and $f_1 = MVN([1,0,0,0], I)$


```{r ls}
mont <- 500 

d <- 4
n <- 20
n0 <- n1 <- n/2

truth <- c(rep(0,n0), rep(1,n1))

mu1 <- rep(0,d)
mu2 <- c(1, rep(0,d-1))
Ls <- 
foreach(i = 1:mont, .combine = 'rbind') %dopar% {
  set.seed(i)
  s0 <- rmvnorm(n0, mean = mu1, sigma = diag(1,d))
  s1 <- rmvnorm(n1, mean = mu2, sigma = diag(1,d))

  samp_dat <- as.matrix(rbind(s0,s1))

  Lrd <- 
    Reduce('cbind',
    lapply(1:d, function(x){
           sum(truth != predict(qda(truth ~ samp_dat[, 1:x]))$class)/length(truth)
  }))
  
  colnames(Lrd) <- paste0("Lr", 1:d)

  as.data.frame(Lrd)
}

```

| measurment | value |
|------------|-------|
| $\mathbb{P}\left[\hat{L}^{(R)}_2 > \hat{L}^{(R)}_1\right]$| <span style="color:red">`r sum(Ls$Lr2 > Ls$Lr1)/nrow(Ls)`</span>|
| $\mathbb{P}\left[\hat{L}^{(R)}_2 = \hat{L}^{(R)}_1\right]$| <span style="color:red">`r sum(Ls$Lr2 == Ls$Lr1)/nrow(Ls)`</span>|
| $\mathbb{P}\left[\hat{L}^{(R)}_3 > \hat{L}^{(R)}_1\right]$| `r sum(Ls$Lr3 > Ls$Lr1)/nrow(Ls)`|
| $\mathbb{P}\left[\hat{L}^{(R)}_3 = \hat{L}^{(R)}_1\right]$| `r sum(Ls$Lr3 == Ls$Lr1)/nrow(Ls)`|
| $\mathbb{P}\left[\hat{L}^{(R)}_4 > \hat{L}^{(R)}_1\right]$| `r sum(Ls$Lr4 > Ls$Lr1)/nrow(Ls)`|
| $\mathbb{P}\left[\hat{L}^{(R)}_4 = \hat{L}^{(R)}_1\right]$| `r sum(Ls$Lr4 == Ls$Lr1)/nrow(Ls)`|


# Box, jitter, and violin

```{r jitter, fig.height = 14, fig.width = 8, echo = FALSE}
m <- melt(LsM)
p <- ggplot(m, aes(x = Var2, y = value)) + xlab("") + ylab("") + geom_hline(yintercept=0, color='red', lty=2)

grid.arrange(
p + geom_boxplot(notch=TRUE, outlier.color=alpha('red',0.5))
, 
p + geom_jitter(pch = 20, height = 0.1, width = 0.3)
,
p + geom_violin(alpha = 0.75)
)

```




